### 2017-05-11 deforestmap.Rmd
Improve discussion

---
title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/conservation-letters.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
---

```{r load_data}
load("deforestmap.rda")
```

#### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A. Rakotomalala<sup>3</sup>, Rija Ranaivosoa<sup>5</sup>, Jean-Roger Rakotoarijaona<sup>5</sup>, Thomas F. Allnutt<sup>4</sup>, and Frédéric Achard<sup>1</sup>

(*) /Email: ghislain.vieilledent@cirad.fr, /Phone: +39.329.457.2273

1. Joint Research Center of the European Commission, Bio-economy Unit (JRC.D.1), I-21027 Ispra (VA), ITALY
2. Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE
3. ETC Terra, F-75016 Paris, FRANCE
4. Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
5. Office National de l'Environnement, 101 Antananarivo, MADAGASCAR

## Summary

1. The island of Madagascar has an unparralleled biodiversity, both in terms of number of species and endemism. This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, which is associated principally to slash-and-burn agriculture and pasture. Scattered data from past studies on deforestation at the national level with substancial cloud cover on available forest maps prevent from analyzing the deforestation trend on the long term and estimate accurate recent annual deforestation rates for Madagascar.

2. In this study, we propose a new approach combining recent (2000-2014) global annual tree cover loss data with historical (1953-2000) national forest-cover maps to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar. With this approach, we obtained cloud-free 30m-resolution forest-cover maps for Madagascar for the year 1990, and annually from 2000 to 2014.

3. We estimated that Madagascar has lost 44% of its natural forest on the period 1953-2014 (37% on the period 1973-2014). Natural forest covered 8.9 Mha in 2014 (15% of the national territory) and was divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangroves. Since 2005, deforestation has progressively increased in Madagascar and has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005. Most of the forest is now located at a distance <1.2km of the forest edge while half of it was located far from forest edge (>1km and up to 4km) in 1973. The percentage of most intact forest as then decreased from 68% in 1973 to 50% in 2014.

4. _Conclusion:_ Despite the conservation efforts in Madagascar, deforestation is dramatically increasing at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions. This increase in the deforestation could be attributable to a rapid population growth (close to 3%/yr) and to a high political instability in the country. More effort, including conservation and development programs, should be done in Madagascar to save the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.

**Keywords**: biodiversity, climate-change, deforestation, Landsat, protected areas, Madagascar, tropical forest, tree cover.

**Potential journals** (2015 IF): Conservation Letters (7.126), Conservation Biology (4.257), Environmental Research Letters (4.134), Biological Conservation (3.985), Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation (2.235), Biotropica (1.944), Tropical Conservation Science (1.55).

## 1. Introduction

Tropical forests provide important ecosystem services at the global scale, such as biodiversity conservation and climate regulation [@Costanza1997] but also many at the local scale for indigeneous populations. At the local scale, tropical forests supply clean water, can preserve populations from flood [@Bradshaw2007] and provide wood for construction and fuel and other non-woody products for food and health. On the island of Madagascar, three types of tropical forests, covering about 9 millions of hectares (15% of the country area) can be found: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016]. Among all the ecosystem services cited previously, Malagasy forests are particularly important for the unique biodiversity they shelter, both in term of number of species and endemism in many taxonomic groups [@Goodman2005]. Deforestation threatens species survival and diminishes biodiversity by destroying forest habitat, creating forest fragments too small to maintain viable populations and increasing 'edge effects' at forest/non-forest interfaces (Harris 1984). Edge effects due to fragmentation typically affect an area several times larger than the forest destruction itself (Harris 1984; Skole & Tucker 1993), affecting micro-meteorology over short distances (Kapos 1989), and increasing exposure to damaging winds (Ferreira & Laurance 1997), fire frequency (Cochrane 2001), and access for livestock, other non-forest animals and hunters (Enserink 1999; Cullen et al. 2000). To assess the impact of tropical deforestation on biological diversity, not only the area deforested, but also the isolation of forest patches and the area of edge habitat must be determined.

Accurate maps of forest-cover and forest-cover change are necessary to monitore carbon emissions and loss of biodiversity associated to deforestation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013a, @Vieilledent2016]. Unfortunately, accurate and cloud-free forest-cover maps are not available for Madagascar for the last decade (2005-2015). In their paper untitled _"Fifty years of deforestation and forest fragmentation in Madagascar"_, @Harper2007 published maps of forest cover change for Madagascar from 1950s to 2000 (for the years _c._ 1953, _c._ 1973, 1990 and 2000). The _c._ 1953 forest map was derived from the visual interpretation of aerial photographies and gives a coarse representation (scale 1/1,000,000) of the forest extent in Madagascar for the 1950s. Forest maps for the years _c._ 1973, 1990 and 2000, were obtained from a supervised classification of Landsat satellite images at 60 (for the year 1973) and 30m (for years 1990 and 2000) and can be used to derive much more accurate estimates of the forest cover at these three dates. Maps provided by @Harper2007 were not free of clouds (11 244 km2 of unknown cover type for the year 2000 for example). Using the same supervised classification approach as in @Harper2007, more recent maps have been released for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009, @ONE2013]. Another recent map using more advanced statistical tool for classification (the Random Forest classifier, see @Grinand2013 and @Rakotomalala2015) has been released for the periods 2005-2010-2013 [@ONE2015]. However, these maps (i) are too old to give recent estimates of deforestation [@MEFT2009, @ONE2013], (ii) figure large areas of clouds (images with more than 20% of clouds in 2010 for the map provided by [@ONE2013]), or (iii) show large errors of classification in specific areas, especially in the dry and spiny forests, for which the spectral answer varies strongly due to deciduousness associated with the higher seasonality than in moist forests (Kappa<0.6 and overall accuracy<0.8 for the dry and spiny forests for the maps produced by @ONE2015). Moreover, deriving such forest maps from supervised classification requires a huge amount of time and work, especially regarding image selection to reduce cloud cover and visual interpretation of the numerous observation polygons used to train the classification algorithm [@Rakotomalala2015]. In addition, visual interpretation might be subjective and is potentially a source of errors. All this work of image selection and visual interpretation have to be repeated to produce new forest maps in the future.

Global forest products have also been published recently and could be used to derive recent forest cover map at the national scale for Madagascar. @Kim2014 produced a global Landsat-based forest-cover change map from 1990 to 2000 but their approach was not reproduced to obtain more recent maps of forest-cover, after 2000. Also, the approach used in @Kim2014 did not allow to correctly map the forest in the dry and spiny ecosystems. Another global study is the work by @Hansen2013 who identified annual tree cover, forest loss and forest gain from 2000 to 2012. Meanwhile, results have been updated and are available up to the year 2014 at http://earthenginepartners.appspot.com/science-2013-global-forest. One problem with using the @Hansen2013 product to map forest cover is that a tree cover threshold must be defined to transform the tree cover data into a forest/non-forest map. Identifying such threshold is not straightforward as forests cannot be defined only by a tree cover value. Also, the tree cover threshold might strongly vary between forest types, mean tree cover values being substantially lower for dry forests than for moist forests. Moreover, a tree cover gives no information on land-use. Therefore, tree plantations such as oil palm or eucalyptus plantations cannot be differenciated from natural forests, which provide very different ecosystem services particularly in terms of biodiversity [@Tropek2014]. Thus, using @Hansen2013 alone can only give a first-approach estimate of the forest cover. But, tree cover is not the only information provided by @Hansen2013. The forest loss, i.e. the complete loss of forest tree cover (which shifts from a positive value over 10% to zero) is also reported.

In this study, we present a simple approach combining maps from @Harper2007 and products from @Hansen2013 to derive annual and cloud-free forest-cover change map for Madagascar for the period 2000-2014. Basically, we used the natural forest-cover map provided by @Harper2007 for the year 2000 (defining the land-use) with the tree cover loss product provided by @Hansen2013 that we applied only inside forest areas identified by @Harper2007. Following the work done by @Harper2007, we studied in details the change in deforestation rates and the evolution of forest fragmentation on six decades from _c._ 1953 to 2014. We showed that despite all the conservation efforts done in Madagascar since the mid 1980s, deforestation has increased in speed since 2005 and that fragmentation has dramatically increased leading to few remaining intact forest landscapes. The approach described in this study could be easilly extended to all tropical countries with a forest map obtained at one date between 2000 and 2014, and could be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated.

## 2. Materials and Methods

### 2.1 Forest-cover maps for Madagascar from _c._ 1953 to 2014

We derived 30m cloud-free annual forest maps for Madagascar for the period 2000-2014 combining the 2000 forest map by @Harper2007 and the 2000-2014 tree cover and tree cover loss maps by @Hansen2013. On the 2000 Harper's forest map, `r as.integer(ha.cloud.2000)` ha of clouds are present, mostly (`r round(perc.cloud.moist)`%) over the moist forest which cover `r round(ha.forest.2000.moist/1e6,2)` million ha. To remove these clouds, we used the 2000 cloud-free tree cover map provided by @Hansen2013 (at 30m resolution) and we chose a threshold of 75% of tree cover [@Achard2014] to decide whether to replace clouds with forest or non-forest. We thus obtained a cloud-free year 2000 forest-cover map for Madagascar. From this map, using tree cover loss data from 2000 to 2014 provided by @Hansen2013, we obtained annual cloud-free forest-cover maps from 2010 to 2014 at 30m resolution. We then corrected the cloudy 1990 Harper forest map assuming that if forest was present in 2000, it was also present in 1990. Doing so, cloud-cover was reduced to a negligeable quantity of `r residual.clouds.1990` ha. For these residual cloud areas, we assumed that forest-cover was the same as for the year 2000. We also corrected the cloudy 1973 Harper forest map assuming that if forest was present in 1990, it was also present in 1973. Contrary to the year 1990, we were not able to remove all the clouds on the 1973 map and the remaining cloud cover was of `r round(cloud.1973/1e6,2)` million ha. Because the 1953 forest-cover map was obtained from the digitalization of a paper map derived from aerial photographies, and thus cannot be perfectly aligned with the maps obtained in the following years, we did not performed any modifications on the original forest map of 1953 except reprojection. For all forest-cover maps from 1973, we removed the isolated non-forest pixels if all the surrounding pixels were covered by forest, assuming that small-scaled deforestation on a 30m square pixel inside the forest was unlikely. This was important in order not to over-estimate forest fragmentation. All these maps are made freely available at https://bioscenemada.cirad.fr/forestmaps.

### 2.2 Computing forest-cover area and deforestation rate

On the basis of the forest-cover maps we obtained, we computed the forest-cover area (in ha) and the mean annual deforestation rate (in ha/yr and in %/yr) on six decades, for five time periods from _c._ 1953 to 2014 (_c._ 1953-_c._ 1973-1990-2000-2010-2014). Mean annual deforestation rates in percentage were computed as follow [@Puyravaud2003; @Vieilledent2013]:

$$\theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]$$

In this formula, $\theta$ is the annual deforestation rate (in %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in years) between the two dates. Because of the large cloud-cover (`r round(cloud.1973/1e6,2)` million ha) in 1973, the annual deforestation rates in hectares of forest for the two periods _c._ 1953-1973 and 1973-1990 are only coarse estimates. This problem was overcome when computing the annual deforestation rates in %/yr for these two periods. In this case, we computed the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the forest pixels that were not covered by clouds in 1973. Computation were done at the national scale and for each forest type in Madagascar: the moist forest in the East, the dry forest in the West, the spiny forest in the South and mangroves on the West coast (Fig. 1). To define the forest type, we used a map provided by the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) showing the boundaries of the four ecoregions in Madagascar. Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996]. Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion on six decades might not be appropriate. As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with caution.

### 2.3 Comparing forest-cover and deforestation rate estimates with previous studies

We compared our estimates of forest-cover and deforestation rate with three previous studies at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015]. @Harper2007 provided forest-cover estimates for the periods _c._ 1953-_c._ 1973-1990-2000, @MEFT2009 provided estimates for the periods 1990-2000-2005 and @ONE2015 provided estimates for the periods 2005-2010-2013. To be able to compare the forest-cover and deforestation rate estimates on the same time-periods, we considered three additional time-periods in our study: 2000-2005, 2005-2010 and 2010-2013. We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies at the same date. All previous studies provide annual deforestation rate estimates in %/yr. The way annual deforestation rates in %/yr have been computed in these previous studies is not always detailed and might slightly differ from one study to another [see @Puyravaud2003]. But these estimates always correct for the potential presences of residuals clouds on forest maps. @Harper2007 also provide deforestation rate estimates in total number of hectares but only for the two periods 1973-1990 and 1990-2000. We converted these estimates into annual deforestation rate estimates in ha/yr. When annual deforestation rate estimates in ha/yr were not reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015), we computed them from the provided forest-cover estimates in each study. These estimates do not correct for the potential presence of residual clouds. Forest-cover and deforestation rates were compared between studies for the whole Madagascar and the four ecoregions. The same ecoregion boundaries as in our study have been used in the @ONE2015 report but this was not the case for the @Harper2007 article and the @MEFT2009 report, which can explain some of the differences between studies in the forest-cover areas and deforestation rates by ecoregions.

### 2.3 Fragmentation

We report an analysis of the evolution of forest fragmentation based on the forest-cover maps for Madagascar for the years _c._ 1953, _c._ 1973, 1990, 2000, 2010 and 2014. We used the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel. Computation were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008]. Six categories of fragmentation are identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined". We used a moving window of 7x7 pixels (4.41 km<sup>2</sup>). Using this window size, forest edge had a width of about 90 m [@Riitters2000]. The interior category can be interpreted as the most intact forest (sensus @Potapov2017). We reported the area of forest in each fragmentation category for the six years and analysed the evolution of the fragmentation on the six decades. We also computed the distance to forest edge (in km) for all forest pixels for the years _c._ 1953, _c._ 1973, 1990, 2000, 2010 and 2014. For that, we used the function `gdal_proximity.py` of the GDAL software (http://www.gdal.org/). We computed the mean and 95% quantiles of the distance to forest edge and looked at the evolution of these values with time.

## 3. Results

### 3.1 Evolution of forest cover and deforestation intensity

Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587041 km<sup>2</sup>. In 2014, this number has dropped down to 8.9 Mha, about 15% of the national territory (Fig. 2 and Tab. 1). Madagascar has then lost 44% of its natural forest in 60 years since c. 1953 and a minimum of 37% in 40 years between c. 1973 and 2014 (Fig. 2 and Tab. 1). In 2014, the 8.9 Mha of natural forest in Madagascar were divided into 4.4 Mha (50%) of moist forest, 2.6 Mha (29%) of dry forest, 1.7 Mha of spiny forest (19%) and 176778 ha (2%) of mangrove forest (Fig. 1 and Tab. 2). Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after 1990 from 204712 ha/yr (1.63%/yr) on the period 1973-1990 to 44296 ha/yr (0.43%/yr) on the period 2000-2005 (Tab. 1). Since 2005, deforestation has progressively increased in Madagascar and deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (Tab. 1). The deforestation trend characterized by a progressive decrease of the deforestation rate on the period 1990-2005 and a progressive increase of the deforestation after 2005 is observable for the four ecoregions (Tab. 3). One exception for the spiny forest for the period 2010-2013 for which we estimated a lower deforestation rate compared to 2005-2010 (Tab. 3).

### 3.2 Comparison with previous forest-cover change studies in Madagascar

Forest-cover maps provided by previous studies [@Harper2007; @MEFT2009; @ONE2015] at the national scale for Madagascar were not free of clouds. In @Harper2007, 5003 and 11244 km2 of unknown cover type were present for the years 1990 and 2000 respectively. Percentage of clouds are not reported in the two other previous studies by @MEFT2009 and @ONE2015. With our approach, we were able to obtain cloud-free forest-cover change maps from 1990 to 2014 (Tab. 1) and thus provide more accurate estimates of forest-cover and deforestation rates on this period. Forest-cover estimates in our study were similar (correlation of 0.99) to estimates from the three previous studies (Tab. 2). We estimated a RMSE of 300000 ha (6% of the mean forest cover of 4.8 Mha) between our forest-cover estimates and estimates from previous studies. A part of these differences can be attributed to differences in ecoregion boundaries between studies. Despite the substantial differences in the deforestation rates estimates between studies (Tab. 3), the same deforestation trend (decrease of the deforestation rate on the period 1990-2005 followed by a progressive increase of the deforestation after 2005) was observed accross studies.

### 3.3 Evolution of forest fragmentation with time

Accompagnying deforestion, forest fragmentation has progressively increased since 1953 in Madagascar. We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014 in Madagascar (Fig. 3). The mean distance to forest edge has decreased to ~300m in 2014 while it was previously of ~1km in 1973 (Fig. 3). Moreover, a large part of the natural forest was located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014 (Fig. 3). The percentage of forest that can be considered intact in Madagascar has progressively decreased since 1953. The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from 68% in 1973 to 50% in 2014 (Tab. 4). In 2014, more than 16% of the forest belonged to the "patch" and "transitional" categories (isolated small forest patches) compared to 9.5% in 1973 (Tab. 4).

## 4. Discussion

### 4.1 Interest of combining recent global annual tree cover loss data with historical national forest-cover maps

In this study, we combined recent (2000-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar. With this approach, we obtained annual cloud-free 30m-resolution forest-cover maps for Madagascar for the period 2000 to 2014. This approach overcome the problem associated to the forest definition which is necessary to transform the 2000 global tree cover data from @Hansen2013 into a forest/non-forest map. Instead, an historical national or regional forest-cover map, based on a more appropriate forest definition at the national or regional scale, can be used as a mask to derive future forest-cover maps until 2014 taking into account global annual tree cover loss data by @Hansen2013. This approach could be easilly extended to other regions or countries with available forest-cover maps at one date on the period 2000-2014. Moreover, that approach can be repeated in the future to obtain future forest-cover change maps, as and when the @Hansen2013 product is updated. The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement). Code should be easilly transposable to other study sites or re-used for future analysis in Madagascar. Of course, the accuracy of the derived forest-cover change maps depends directly on the accuracy of the historical forest-cover maps and the accuracy of the tree cover loss data. Accuracy of the tree cover loss data [@Hansen2013] has been discussed in several recent studies [@Verhegghen2016].

Regarding the potential forest regrowth, we made the same choice as @Harper2007 and did not consider it in our study, although @Hansen2013 are also providing tree cover gain data over the period 2010-2014. We made this choice for several reasons. First, there is little evidence of natural forest regeneration in Madagascar on the long term [@Harper2007, @Grouzis2001]. This can be explained by several ecological mechanisms following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001]. Moreover, in the areas where forest regeneration is ecologically possible, young regrowth forests are more easilly re-burnt for agriculture and pasture. Finally, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.

### 4.2 Evolution of forest-cover in Madagascar since 1953 and current forest-cover

We estimated that natural forests covered 8.9 Mha in Madagascar (15% of the national territory of 587041 km<sup>2</sup>) in 2014 and that Madagascar has lost 44% of its natural forest in six decades since 1953 (37% since 1973). There is an age-old debate about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscape since their large-scale settlement around 800 CE [@Cox2012]. French naturalists Henri Humbert [@Humbert1927] and Henri Perrier de la Bâthie [@Perrier1921] stated that the full island was originally covered by forest, leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000]. More recent studies counter-balanced that point of view saying that grasslands existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017]. Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000]. Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the forest landscape and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years. Associated to deforestation, we showed in our study that residual forest in Madagascar were highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014. Both deforestation and forest fragmentation has strong consequences on biodiversity and carbon emissions in Madagascar. Around 90% of Madagascar's endemic species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 had led to an extinction of 9% of the species.

Increased wind speed, vapour pressure deficit and disturbance pressure in the vicinity of forest edges increase tree mortality, resulting in persistent carbon emissions long after deforestation occurs. Fire, wind and desiccation can penetrate up to kilometres into the forest15,19,20, altering microclimate within 100-300m of the forest edge13,21,22,23. Field studies have estimated relative carbon losses in forest edges (e) averaging 11% (ref. 24) and up to 36% (ref. 22) and 50% (ref. 25) of aboveground biomass within the 100m edge zone, and modelling studies have shown biomass reductions of up to 70% in small fragments2

Increase in forest fragmentation has a great impact on both carbon emissions [@Putz2014] and biodiversity loss [@Gibson2013, @Murcia1995].
Biodiversity loss, see @Allnutt2008.

More efforts need to be done to save the few remaining natural tropical forests.

### 4.3 Deforestation trend and causes of the acceleration of the deforestation since 2005

In our study, we have shown that the progressive decrease of the deforestation rate on the period 1990-2005 was followed by a continuous increase in the deforestation rate on the period 2005-2014. We showed that deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (42296 ha.yr, 0.43%/yr). These results are confirmed by previous studies on deforestation at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015]. Despite differences in the methodologies regarding (i) forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs. Random Forests for @ONE2015), (iii) deforestation rate computation, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar. The increase of the deforestation rates after 2005 could be attributed to both population growth and political instability in the country. Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013, @Gorenflo2011]. With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in the deforestation rate. The political crisis of 2009, followed by several years of political instability and weak governance is also very likely to explain the increase in the deforestation rate observed on the period 2005-2014. Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from the previous studies, in particular for deforestation rates expressed in number of hectares per year. Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015]. Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr, if we consider the period 2010-2014) could be used to build reference scenarios for deforestation in Madagascar [@Olander2008].

## 5. Tables

```{r defor, results="asis"}
defor.df <- read.table(file="outputs/defor.txt", sep="\t", header=TRUE)
perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
perc.2 <- round(100*(defor.df$area[defor.df$Year==1953]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1953])
d.1 <- defor.df$theta[defor.df$Year==1990]
d.2 <- defor.df$theta[defor.df$Year==2010]
tab.1 <- defor.df
names(tab.1) <- c("Year", "Forest (ha)", "Clouds (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
kable(tab.1)
```

Table 1: **Evolution of the forest cover and deforestation rates from 1950s to 2014 in Madagascar**. Madagascar has lost `r perc.1`% of its natural forest in 40 years between _c._ 1973 and 2014 (and about `r perc.2`% in 60 years since _c._ 1953). Deforestation has slew down after 1990 (from `r d.1` in 1973-1990 to `r d.2`%/yr in 2000-2010), most likely due to political changes (end of the Second Republic of Madagascar in 1992), a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility. But more recently, deforestation has increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010. This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014). Natural forests cover 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70000 ha/yr.

```{r fcc_comp, results="asis"}
fcc.comp <- read.table(file="outputs/fcc_comp.txt", sep=" ", header=TRUE)
tab.2 <- fcc.comp
tab.2 <- tab.2[,c(2,1,3:10)]
tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
kable(tab.2,col.names=col.names,align=c("l","l",rep("r",8)))
#print(xtable(tab.2,col.names=col.names,align=c("llrrrrrrrrr")),type="html")
```

Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**. We found forest-cover estimates similar to previous studies [@Harper2007;@MEFT2009;@ONE2015]. The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map. Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies. Our method provides cloud-free forest-cover estimates from 1990 to 2014 and corrections for the forest-cover in 1973.

```{r defor_comp, results="asis"}
tab.3 <- defor.comp
tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
kable(tab.3,col.names=col.names,align=c("l","l",rep("r",6)))
```

Table 3: **Comparison of Madagascar annual deforestation rates with previous studies on the period 1953-2014**. Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis). For deforestation rate in %/yr, exact same numbers as in scientific articles and reports from previous studies have been reported. The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another but estimates always correct for the potential presences of clouds on forest maps. Deforestation rates in ha/yr have been recomputed from forest-cover estimates in Tab. 2 (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforestation rates in ha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.

```{r frag, results="asis"}
frag.df <- read.table(file="outputs/frag.txt", sep="\t", header=TRUE)
tab.4 <- frag.df
p.1 <- round(frag.df$interior[frag.df$Year==1973])
p.2 <- round(frag.df$interior[frag.df$Year==2014])
p.3 <- round(frag.df$patch[frag.df$Year==2014]+frag.df$transitional[frag.df$Year==2014])
names(tab.4) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
kable(tab.4)
```

Table 4: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**. Forest fragmentation has progressively increased since 1953. The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in 1973 to `r p.2`% in 2014. More than `r p.3`% of the forest belong to the "patch" and "transitional" categories (isolated small forest patches) in 2014.

## 6. Figures

![Figure 1: **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue). Ecoregions were defined following climatic (see @Cornet1974) and vegetation (see @IEFN1996) criteria. The dark grey areas represent the remaining natural forest cover for the year 2014.](outputs/ecoregion.png)

![Figure 2: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset. Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014. Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](outputs/fig_fcc_highres.png)

![Figure 2: **Evolution of the mean distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year. Vertical dashed segments represent the 95% quantiles of the distance to forest edge. We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014. The mean distance was of ~1km in 1973 while it was of ~300m in 2014. Moreover, many areas were located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014.](outputs/dist.png)

## Acknowledgements

This study is part of the Cirad's BioSceneMada project (https://bioscenemada.cirad.fr) and the Joint Research Center's ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd). The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD project is funded by the European Commission.

## Data availability statement

All the data and codes used for this study are made publicly available in the `deforestmap` GitHub repository (https://github.com/ghislainv/deforestmap.git). The results and the manuscript are fully reproducible running the R script `deforestmap.R` located inside the `deforestmap` repository.

## References