### 2017-05-29 deforestmap.Rmd
Comments by Fred

---
title: "Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar"
bibliography: bib/biblio.bib
csl: bib/biological-conservation.csl
output: 
  html_document:
    fig_caption: yes
  pdf_document:
    fig_caption: true
    keep_tex: true
  word_document:
    fig_caption: yes
---

```{r load_data}
load("deforestmap.rda")
```

#### Ghislain Vieilledent<sup>*,1,2</sup>, Clovis Grinand<sup>3</sup>, Fety A. Rakotomalala<sup>3</sup>, Rija Ranaivosoa<sup>5</sup>, Jean-Roger Rakotoarijaona<sup>5</sup>, Thomas F. Allnutt<sup>4</sup>, and Frédéric Achard<sup>1</sup>

(*) /Email: ghislain.vieilledent@cirad.fr, /Phone: +39.329.457.2273

1. Joint Research Center of the European Commission, Bio-economy Unit (JRC.D.1), I-21027 Ispra (VA), ITALY
2. Cirad, UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE
3. ETC Terra, F-75016 Paris, FRANCE
4. Wildlife Conservation Society, 101 Antananarivo, MADAGASCAR
5. Office National pour l'Environnement, 101 Antananarivo, MADAGASCAR

## Summary

1. _Context:_ The island of Madagascar has an unparalleled biodiversity, both in terms of number of species and endemism. This biodiversity is mainly located in the tropical forests of the island and is highly threatened by anthropogenic deforestation, which is associated principally to slash-and-burn agriculture and pasture. Scattered forest maps from past studies at national level with substantial gaps (due to presence of cloud cover on satellite imagery) prevent from analyzing the deforestation trends on a long term perspective in Madagascar.

2. _Approach:_ In this study, we propose a new approach combining historical (1953-2000) national forest-cover maps with recent (2000-2014) global annual tree cover loss data to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar. We also produce new forest-cover maps at 30m-resolution over the full territory of Madagascar for the year 1990, and annually from 2000 to 2014.

3. _Results:_ We estimate that Madagascar has lost 44% of its natural forest cover over the period 1953-2014 (including 37% over the period 1973-2014). Natural forests cover 8.9 Mha in 2014 (15% of the national territory) which are divided into 4.4 Mha (50%) of moist forests, 2.6 Mha (29%) of dry forests, 1.7 Mha of spiny forests (19%) and 177,000 ha (2%) of mangroves. Since 2005, the annual deforestation rate has progressively increased in Madagascar and has more than doubled from the period 2000-2005 to reach 99,000 ha/yr  during 2010-2014 (corresponding to a rate of 1.08%/yr). Most of the forests are now located at less than 1.2km from forest edges while half of them were located far from forest edges in 1973  (i.e. between 1km and 4km from forest edges). The percentage of core forest has then decreased from 68% in 1973 to 50% in 2014.

4. _Conclusion:_ Despite the conservation efforts in Madagascar, deforestation has continued to significantly increase at the national level since 2005, with strong consequences on biodiversity loss and carbon emissions. This increase in the deforestation rate can be related to a rapid population growth (close to 3%/yr) and to poor law enforcement due to political instability in the country. More effort, including conservation and development programs, should be done in Madagascar to protect the remaining natural tropical forests, both to enhance local people livelihoods and answer to global environmental problems.

**Keywords**: biodiversity, climate-change, deforestation, Landsat, protected areas, Madagascar, tropical forest, tree cover.

**Potential journals** (2015 IF): Conservation Letters (7.126), Conservation Biology (4.257), Environmental Research Letters (4.134), Biological Conservation (3.985), Landscape Ecology (3.657), Ambio (2.555), Environmental Conservation (2.235), Biotropica (1.944), Tropical Conservation Science (1.55).

## 1. Introduction

Separated from the African continent and the Indian plate about 165 and 88 million years ago respectively [@Ali2008], the flora and fauna of Madagascar followed its own evolutionary path. Combined with a high number of micro-habitats [@Pearson2009], this has led to Madagascar's exceptional biodiversity both in term of number of species and endemism in many taxonomic groups [@Crottini2012; @Goodman2005]. Most of the biodiversity in Madagascar is concentrated in the tropical forests of the island which can be divided in three types: the moist forest in the East, the dry forest in the West and the spiny forest in the South [@Vieilledent2016]. This unparalleled biodiversity is severely threatened by anthropogenic deforestation [@Harper2007; @Vieilledent2013] associated to activities such as slash-and-burn agriculture and pasture [@Scales2011]. Tropical forests in Madagascar also store a large amount of carbon [@Vieilledent2016] and high rates of deforestation in Madagascar are responsible for large CO2 emissions in the atmosphere [@Achard2014]. Deforestation threatens species survival by reducing their habitat [@Brooks2002; @Tidd2001]. Associated to deforestation, forest fragmentation can also lead to species extinction by isolating populations from each others and creating forest patches too small to maintain viable populations [@Saunders1991]. Fragmentation also extents forest edge where ecological conditions (such as air temperature, light intensity and air moisture) can be dramatically modified, with consequences on the abundance and distribution of species [@Murcia1995]. Forest fragmentation can also have substantial effects on forest carbon storage capacity, carbon stocks being much lower at the forest edge than under a closed canopy [@Brinck2017]. Moreover, forest carbon stocks are varying spatially due to climate or soil factors [@Saatchi2011; @Vieilledent2016]. As a consequence, accurate and spatially explicit maps of forest-cover and forest-cover change are necessary to monitor loss of biodiversity and carbon emissions associated to deforestation and forest fragmentation, assess the efficiency of present conservation strategies [@Eklund2016], and implement new strategies for the future [@Vieilledent2013; @Vieilledent2016]. Simple time-series of forest-cover estimates, such as those provided by the FAO Forest Resource Assessment report [@Keenan2015] are not sufficient.

Unfortunately, accurate and exhaustive forest-cover maps are not available for Madagascar for the last fifteen years (2000-2015). @Harper2007 produced maps of forest cover and forest cover changes over Madagascar for the years c. 1953, c. 1973, 1990 and 2000. The c. 1953 forest map was derived from the visual interpretation of aerial photography at coarse scale (1/1,000,000). Forest maps for the years c. 1973, 1990 and 2000, were obtained from supervised classification of Landsat satellite images at 60m resolution (for the year 1973) or 30m resolution (for years 1990 and 2000) and can be used to derive accurate estimates (89.5% accuracy reported for the forest / non-forest map of year 2000). Maps provided by @Harper2007 were not exhaustive (due to the presence of clouds in the satellite imagery), e.g. 11 244 km2 are mapped as unknown cover type for the year 2000. Using a similar supervised classification approach as in @Harper2007, more recent maps were produced for the periods 2000-2005-2010 by national institutions, with the technical support of international environmental NGOs [@MEFT2009; @ONE2013]. Another set of recent forest cover maps using an advanced statistical tool for classification, the Random Forest classifier [@Grinand2013; @Rakotomalala2015], was produced for the periods 2005-2010-2013 [@ONE2015]. However, these maps are either too old to give recent estimates of deforestation [@MEFT2009; @ONE2013], include large areas of missing information (due to images with more than 20% of clouds) for the 2010 map provided by [@ONE2013], or show large mis-classification in specific areas, especially in the dry and spiny forest domain for which the spectral answer has a strong seasonal behavior due to the deciduousness of such forests (overall accuracy is lower than 0.8 for the dry and spiny forests for the maps produced by @ONE2015). Moreover, the production of such forest maps from a supervised classification approach requires a significant amount of resources, especially regarding the image selection step (required to minimize cloud cover) and the training step (visual interpretation of a large number of polygons needed to train the classification algorithm) [@Rakotomalala2015]. Most of this work of image selection and visual interpretation would need to be repeated if one would intend to produce new forest maps in the future using a similar approach.

Global forest or tree cover products were also published recently and can be tested at the national scale for Madagascar. [@Kim2014] produced a global forest-cover change map from 1990 to 2000 (derived from Landsat imagery). This product was updated to cover the period 1975-2005 (<http://glcf.umd.edu/data/landsatFCC/>) but forest-cover maps after 2005 were not produced. Moreover, the approach used in @Kim2014 did not allow to correctly map the forests in the dry and spiny ecosystems of Madagascar (see Fig. 8 in @Kim2014). @Hansen2013 mapped tree cover percentage, annual forest loss and forest gain from 2000 to 2012 at global scale and at 30 m resolution. Meanwhile, this product was updated and is now available up to the year 2014 [@Hansen et al. 2013]. To map forest cover from the @Hansen2013 product, a tree cover threshold must be selected (to define forest cover). Selecting such a threshold is not straightforward as the accuracy of the global tree cover map strongly varies between forest types, being substantially lower for dry forests than for moist forests [@Bastin2017]. Moreover, the @Hansen2013 product does not provide information on land-use. In particular the global tree cover map does not separate tree plantations such as oil palm or eucalyptus plantations from natural forests [@Tropek2014]. Thus, the global tree cover map from @Hansen2013 cannot be used alone to produce a map of forest cover [@Tuykavina2017]. In complement to the tree cover percentage provided in @Hansen2013, a layer of annual forest loss is also provided (i.e. complete loss of tree cover from a value higher than 10% to zero).

In this study, we present a simple approach which combines the maps from @Harper2007 and products from @Hansen2013 to derive annual wall-to-wall forest-cover change maps over the period 2000-2014 for Madagascar. Similarly to the approach of @Harper2007, we assess then in details the trends in deforestation rates and forest fragmentation from c. 1953 to 2014. The approach described in this study should help assessing the efficiency of present conservation strategies and implementing new strategies for the future. Our approach could be easily extended to other tropical countries when at least a forest cover map exist between 2000 and 2014 or repeated in the future as and when the @Hansen2013 products are updated.

## 2. Materials and Methods

### 2.1 Creation of new forest-cover maps of Madagascar from 1953 to 2014

We produced annual forest/non-forest maps at 30m resolution for the full territory of Madagascar for the period 2000-2014 by combining the forest map of year 2000 from @Harper2007 and the tree cover percentage and annual forest cover loss maps over the period 2000-2014 from @Hansen2013. The 2000 Harper's forest map includes 208,000 ha of unclassified areas due to the presence of clouds in input satellite imagery, mostly (88%) within the moist forest domain which covers 4.17 million ha in total. To provide a label (forest or non-forest) to these unclassified pixels, we use the 2000 tree cover percentage map of @Hansen2013 by selecting a threshold of 75% tree cover to define forest cover as recommended by other studies for the moist domain [@Achard2014]. We thus obtained a forest-cover map for the year 2000 covering the full territory of Madagascar. We then combined this forest cover map of the year 2000 with the annual forest cover loss maps from 2001 to 2014 provided by @Hansen2013 to create annual forest cover maps from 2001 to 2014 at 30m resolution. In a further step we completed the Harper’s forest map of year 1990 by filling unclassified areas using our forest cover map of year 2000: we assume that if forest was present in 2000, the pixel was also forested in 1990. The remaining unclassified pixels were limited to a relatively small total area of c. 8,000 ha. We labeled these residual pixels as non-forest as for the year 2000. Similarly we completed the Harper forest map of year 1973 by filling unclassified areas using our forest cover map of the year 1990 assuming that if forest was present in 1990, it was also present in 1973. Contrary to the year 1990, the remaining unclassified pixels for year 1973 corresponded to a significant total area of 3.32 million ha, which could not be labeled. The forest-cover map of year 1953 was only re-projected to a common projection. This map was produced by scanning a paper map derived from aerial photos, and thus could not be perfectly aligned with the other maps produced through digital processing of satellite imagery. Finally for all forest-cover maps from 1973, the isolated single non-forest pixels (i.e. fully surrounded by forest pixels) were removed, assuming that single non-forest pixels inside a forest patch were not corresponding to deforestation (they might correspond to selective logging activities). This allowed avoiding to include very small scale events (<0.1 ha such as selective logging) as forest fragmentation. All the resulting maps are freely available at <https://bioscenemada.cirad.fr/forestmaps>.

### 2.2 Computing forest-cover areas and deforestation rates

From these new forest-cover maps  we computed the total forest-cover area for six available years (1953-1973-1990-2000-2010-2014), and the annual deforested area and annual deforestation rate for the corresponding five time periods between 1953 and 2014. The annual deforestation rates were computed as follow [@Puyravaud2003; @Vieilledent2013]:

$$\theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}]$$

where $\theta$ is the annual deforestation rate (in %/yr), $F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in years) between the two dates.

Because of the large unclassified area (3.32 million ha) in 1973, the annual deforestation areas and rates for the two periods 1953-1973 and 1973-1990 are only indicative estimates. For these two periods the annual deforestation rates are computed as the ratio $(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels. Area and rate estimates are produced at the national scale and for the four forest ecosystems present in Madagascar: moist forest in the East, dry forest in the West, spiny forest in the South, and mangroves on the Western coast (Fig. 1). To define the forest domains, we used a map from the MEFT (_"Ministère de l'Environnement et des Forêts à Madagascar"_) with the boundaries of the four ecoregions in Madagascar. Ecoregions were defined on the basis of climatic and vegetation criteria using the climate classification by @Cornet1974 and the vegetation classification from the 1996 IEFN national forest inventory [@IEFN1996]. Because mangrove forests are highly dynamic ecosystems that can expand or contract on decadal scales depending on changes in environmental factors [@Armitage2015], a fixed delimitation of the mangrove ecoregion on six decades might not be fully appropriate. As a consequence, our estimates of the forest-cover and deforestation rates for mangroves in Madagascar must be considered with this limitation.

### 2.3 Comparing our forest-cover and deforestation rate estimates with previous studies

We compared our estimates of forest-cover and deforestation rates with estimates from the three existing studies at the national scale for Madagascar : (i) [@Harper2007], (ii) [@MEFT2009] and (iii) [@ONE2015]. @Harper2007 provides forest-cover and deforestation estimates for the periods c. 1953-c. 1973-1990-2000. MEFT, USAID, and CI (2009) provides estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and Etc Terra (2015) provides estimates for the periods 2005-2010-2013. To compare our forest-cover and deforestation estimates over the same time periods, we consider three additional time-periods in our study (2000-2005, 2005-2010 and 2010-2013) by creating extra forest cover maps for years 2005 and 2013. We computed the Pearson's correlation coefficient and the root mean square error (RMSE) between our forest-cover estimates and forest-cover estimates from previous studies for all the dates and the four forest types. For previous studies, the computation of annual deforestation rates (in %/yr) is not always detailed and might slightly differ from one study to another [see @Puyravaud2003]. @Harper2007 also provides total deforested areas for the two periods 1973-1990 and 1990-2000. We converted these values into annual deforested area estimates. When annual deforested areas were not reported (for 1953-1973 in @Harper2007 and in @MEFT2009 and @ONE2015), we computed them from the forest-cover estimates in each study. These estimates cannot be corrected from the potential bias due to the presence of residual clouds. Forest-cover and deforestation rates were then compared between all studies for the whole Madagascar and the four ecoregions. The same ecoregion boundaries as in our study were used in @ONE2015 but this was not the case for @Harper2007 and @MEFT2009, which can explain part of the differences between the estimates.

### 2.3 Fragmentation

An analysis of the changes in forest fragmentation was carried out from our forest-cover maps over Madagascar for the years 1953, 1973, 1990, 2000, 2010 and 2014. We applied the method developed by @Riitters2000 which uses a moving window to characterize the fragmentation around each forested pixel. Computation were done using the function `r.forestfrag` of the GRASS GIS software [@Neteler2008]. Six categories of fragmentation were identified from the amount of forest and its occurrence as adjacent forest pixels: "interior", "perforated", "edge", "transitional", "patch", and "undetermined". We used a moving window of 7x7 pixels (4.4 ha). Using this window size, forest edge had a width of about 90m [@Riitters2000]. The interior category can be interpreted as the most intact forest [@Potapov2017]. We reported the area of forest in each fragmentation category for the six years and analyzed the dynamics of fragmentation over the six decades. We also computed the distance to forest edge for all forest pixels for the years 1953, 1973, 1990, 2000, 2010 and 2014. For that, we used the function `gdal_proximity.py` of the GDAL software (http://www.gdal.org/). We computed the mean and 95% quantiles of the distance to forest edge and looked at the evolution of these values with time.

## 3. Results

### 3.1 Dynamics of forest cover and deforestation intensity

Natural forests in Madagascar covered 16.0 Mha in 1953, about 27% of the national territory of 587,041 km2. In 2014, the forest cover has dropped down to 8.9 Mha, corresponding to about 15% of the national territory (Fig. 2 and Tab. 1). Madagascar has lost 44% and 37% of its natural forests between 1953 and 2013, and between c. 1973 and 2014 respectively (Fig. 2 and Tab. 1). In 2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4 Mha of moist forest (50% of total forest cover), 2.6 Mha of dry forest (29%), 1.7 Mha of spiny forest (19%) and 0.18 Mha (2%) of mangrove forest (Fig. 1 and Tab. 2). Regarding the deforestation trend, we observed a progressive decrease of the deforestation rate after1990 from 205,000 ha/yr (1.63%/yr) over the period 1973-1990 to 44,300 ha/yr (0.43%/yr) over the period 2000-2005 (Tab. 1). Then from 2005, the deforestation rate has progressively increased and has more than doubled over the period 2010-2014 (98,700 ha/yr, 1.08%/yr) compared to 2000-2005 (Tab. 1). The deforestation trend characterized by a progressive decrease of the deforestation rate over the period 1990-2005 and a progressive increase of the deforestation after 2005 is valid for all four ecoregions (Tab. 3), excepting the spiny forest domain for which the deforestation rate during the period 2010-2013 was lower than during 2005-2010 (Tab. 3).

### 3.2 Comparison with previous forest-cover change studies in Madagascar

Forest-cover maps provided by previous studies over Madagascar were not exhaustive (unclassified areas) due to the presence of clouds in satellite imagery used to produce such maps. In @Harper2007, the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown cover type respectively. Proportions of unclassified areas are not reported in the two other existing studies by @MEFT2009 and @ONE2015. With our approach, we produced wall to wall forest-cover change maps from 1990 to 2014 for the full territory of Madagascar (Tab. 1). This allows to produce more robust estimates of forest-cover and deforestation rates over this period. Quite logically, our forest-cover estimates over the period 1953-2013 (considering forest cover estimates at national level and by ecoregions for the 6 available dates) were well correlated  (0.99) to estimates from the three previous studies (Tab. 2) with a RMSE of 300,000 ha (6% of the mean forest cover of 4.8 Mha when considering all dates and forest types together). These small differences can be partly  attributed to differences in ecoregion boundaries. Despite significant differences in deforestation estimates (Tab. 3), a similar deforestation trend was observed across studies with a decrease of deforestation rates over the period 1990-2005 followed by a progressive increase of the deforestation after 2005.

### 3.3 Evolution of forest fragmentation with time

In parallel to the dynamics of deforestation, forest fragmentation has progressively increased since 1953 in Madagascar. We observed a continuous decrease of the mean distance to forest edges from 1953 to 2014 in Madagascar. The mean distance to forest edge has decreased to c. 300m in 2014 while it was previously c. 1km in 1973 (Fig. 3). Moreover, a large proportion of the natural forests were located far from forest edges in 1973 (between 2km and 4km), while the majority of forests were at less than 1.2 km from forest edges in 2014 (Fig. 3). The percentage of forest that can be considered intact in Madagascar has continuously decreased since 1953: the percentage of forest belonging to the "interior" category (most intact forests) has fallen from 68% in 1973 to 50% in 2014. In 2014, more than 16% of the forest belonged to the "patch" and "transitional" categories (isolated small forest patches) compared to 9.5% in 1973 (Tab. 4).

## 4. Discussion

### 4.1 Interest of combining recent global annual tree cover loss data with historical national forest-cover maps

In this study, we combined recent (2001-2014) global annual tree cover loss data [@Hansen2013] with historical (1953-2000) national forest-cover maps [@Harper2007] to look at six decades (1953-2014) of deforestation and forest fragmentation in Madagascar. We obtained annual cloud-free 30m-resolution forest-cover maps for Madagascar for the period 2000 to 2014. This approach overcomes the problem associated to the forest definition which is necessary to transform the 2000 global tree cover data by @Hansen2013 into a forest/non-forest map [@Tropek2014]. Instead, an historical national or regional forest-cover map, based on a more appropriate forest definition at the national or regional scale, can be used as a mask. This approach could be easilly extended to other regions or countries with available forest-cover maps at one date in the period 2000-2014. Moreover, that approach can be repeated in the future to obtain future forest-cover maps, as and when the @Hansen2013 product is updated. The R/GRASS code used for this study is made freely available in a GitHub repository (see Data availability statement) and should be easilly transposable to other study sites or re-used for future analysis in Madagascar. Of course, the accuracy of the derived forest-cover change maps depends directly on the accuracy of the historical forest-cover maps and the accuracy of the tree cover loss data. Reported global accuracy for tree cover loss data on the period 2000-2012 is 99.6% (see Tab. S5 in @Hansen2013). @Verhegghen2016 also found similar annual deforestation rate estimates when comparing results derived from the tree cover loss data by @Hansen2013 on the period 2000-2012 with results derived from semi-supervized classification of Landsat satellite images on the period 2000-2010 by @Achard2014 for six countries in Central Africa. Consistently with the study by @Harper2007, we did not consider the potential natural forest regrowth in Madagascar, although @Hansen2013 are also providing tree cover gain data over the period 2001-2014. We made this choice for several reasons. First, gain data provided by @Hansen2013 can figure plantations and not natural forests [@Tropek2014]. Second, there is little evidence of natural forest regeneration in Madagascar on the long term [@Harper2007, @Grouzis2001]. This can be explained by several ecological mechanisms following burning practice such as soil erosion [@Grinand2017] and reduced seed bank due to fire and soil loss [@Grouzis2001]. Moreover, in the areas where forest regeneration is ecologically possible, young regrowth forests are more easilly re-burnt for agriculture and pasture. Third, young secondary forests do not provide the same ecosystem services (in terms of biodiversity and carbon storage) as old-growth natural forests.

### 4.2 Evolution of forest-cover in Madagascar since 1953 and current forest-cover

We estimated that natural forests covered 8.9 Mha in Madagascar (15% of the national territory of 587041 km<sup>2</sup>) in 2014 and that Madagascar has lost 44% of its natural forest in six decades since 1953 (37% since 1973). There is an age-old debate [@Burns2016] about the original forest cover in Madagascar to know to which extent humans have altered the natural forest landscape since their large-scale settlement around 800 CE [@Cox2012]. French naturalists Henri Humbert [@Humbert1927] and Henri Perrier de la Bâthie [@Perrier1921] stated that the full island was originally covered by forest, leading to the common statement that 90% of the natural forests have disappeared since the arrival of humans on the island [@Kull2000]. More recent studies counter-balanced that point of view saying that grasslands existed in Madagascar long before human arrival and were determined by climate, natural grazing and other natural factors [@Virah-Sawmy2009, @Vorontsova2017]. Other authors have questioned the entire narrative of extensive alteration of the landscape by early human activity which, through legislation, has severe consequences on local people [@Klein2002; @Kull2000]. Whatever the original proportion of natural forests and grasslands in Madagascar, our results demonstrate that human activities since the 5Os have profoundly impacted the natural tropical forests and that recent conservation and development programs in Madagascar have failed to stop deforestation in the recent years. Deforestation has strong consequences on biodiversity and carbon emissions in Madagascar. Around 90% of Madagascar's species are forest dependent [@Allnutt2008; @Goodman2005] and @Allnutt2008 estimated that deforestation between 1953 and 2000 had led to an extinction of 9% of the species. The additional deforestation we observed on the period 2000-2014 (around 1Mha of deforested natural forest) worsen this result. Regarding carbon emissions, using the 2010 aboveground forest carbon map by @Vieilledent2016, we estimated that deforestation on the period 2010-2014 has led to the emissions of 40224 Gg (1 Gg = 10<sup>9</sup> g) of carbon in the atmosphere (10055 Gg/yr) and that the residual aboveground forest carbon stock was of 832822 Gg in 2014. Associated to deforestation, we showed that residual forest in Madagascar was highly fragmented with 95% of the forest being at a distance <1.2 km of the forest edge in 2014. Forest fragmentation creates forest fragments too small to maintain viable populations and increases 'edge effects' at forest/non-forest interfaces having great impact on both carbon emissions [@Brinck2017] and biodiversity loss [@Gibson2013; @Murcia1995].

### 4.3 Deforestation trend and causes of the acceleration of the deforestation since 2005

In our study, we have shown that the progressive decrease of the deforestation rate on the period 1990-2005 was followed by a continuous increase in the deforestation rate on the period 2005-2014. We showed that deforestation rate has more than doubled on the period 2010-2014 (98651 ha/yr, 1.08%/yr) compared to 2000-2005 (42296 ha.yr, 0.43%/yr). These results are confirmed by previous studies on deforestation at the national scale for Madagascar [@Harper2007; @MEFT2009; @ONE2015]. Despite differences in the methodologies regarding (i) forest definition (associated to independant visual interpretations of observation polygons to train the classifier), (ii) classification algorithms (Maximum Likelihood Classification for @Harper2007 and @MEFT2009 vs. Random Forests for @ONE2015), (iii) deforestation rate computation method, and (iii) correction for the presence of clouds, these studies gave similar forest-cover estimates and show the same deforestation trend over Madagascar. Our deforestation rate estimates from 1990 to 2014 have been computed from cloud-free maps and can be considered more accurate in comparison with estimates from these previous studies, in particular for the deforestation rates expressed in number of hectares per year. Forest-cover and deforestation rate estimates from this study could be used to update the estimates in the upcoming FAO Forest Resources Assessment reports [@Keenan2015]. Current rates of deforestation for Madagascar (98651 ha/yr, 1.08 %/yr for 2010-2014) could also be used to build reference scenarios for deforestation in Madagascar and help implement deforestation mitigation initatives such as the REDD+ [@Olander2008]. The increase of the deforestation rates we observed after 2005 could be attributed to both population growth and political instability in the country. Nearly 90% of Madagascar's population relies on biomass for their daily energy needs [@Minten2013] and the link between population size and deforestation has previously been demonstrated in Madagascar [@Vieilledent2013; @Gorenflo2011]. With a mean demographic growth rate of about 2.8%/yr and a population which has increased from 16 to 24 million people on the period 2000-2015 [@UN2015], the increasing demand in wood-fuel and space for agriculture is likely to explain the increase in deforestation rates. The political crisis of 2009 [@Ploch2012], followed by several years of political instability and weak governance could also explain the increase in the deforestation rate observed on the period 2005-2014 [@Smith2003]. These results show that despite the conservation policy in Madagascar [@Freudenberger2010], deforestation has dramatically increased at the national level since 2005. More efforts should be done to identify the real causes and mechanisms explaining this rampant deforestation in order to implement new conservation strategies that will be effective at conserving the few remaining natural tropical forests in Madagascar.

## 5. Tables

```{r defor, results="asis"}
defor.df <- read.table(file="outputs/defor.txt", sep="\t", header=TRUE)
perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
perc.2 <- round(100*(defor.df$area[defor.df$Year==1953]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1953])
d.1 <- defor.df$theta[defor.df$Year==1990]
d.2 <- defor.df$theta[defor.df$Year==2010]
tab.1 <- defor.df
names(tab.1) <- c("Year", "Forest (ha)", "Clouds (ha)", "Annual defor. (ha/yr)", "theta (%/yr)")
kable(tab.1)
```

Table 1: **Evolution of the forest cover and deforestation rates from 1950s to 2014 in Madagascar**. Madagascar has lost `r perc.1`% of its natural forest in 40 years between _c._ 1973 and 2014 (and about `r perc.2`% in 60 years since _c._ 1953). Deforestation has slew down after 1990 (from `r d.1` in 1973-1990 to `r d.2`%/yr in 2000-2010), most likely due to political changes (end of the Second Republic of Madagascar in 1992), a decrease of available arable lands under the remaining forests and to a decrease of forest accessibility. But more recently, deforestation has increased, with a doubling deforestation on the period 2010-2014 compared to 2000-2010. This increase could be attributed to both political instability (political crisis in 2009) and to an explosion of the demography (according to the @UN2015, population has increased from about 16M inhabitants in 2000 to 24M in 2014). Natural forests cover 8.9 Mha in 2014 and current (period 2000-2014) mean annual deforestation rate is of about 70000 ha/yr.

```{r fcc_comp, results="asis"}
fcc.comp <- read.table(file="outputs/fcc_comp.txt", sep=" ", header=TRUE)
tab.2 <- fcc.comp
tab.2 <- tab.2[,c(2,1,3:10)]
tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2014")
kable(tab.2,col.names=col.names,align=c("l","l",rep("r",8)))
#print(xtable(tab.2,col.names=col.names,align=c("llrrrrrrrrr")),type="html")
```

Table 2: **Comparison of Madagascar forest-cover estimates (in ha) with previous studies on the period 1953-2014**. We found forest-cover estimates similar to previous studies [@Harper2007;@MEFT2009;@ONE2015]. The increase in mangrove and spiny forest covers from _c._ 1953 to _c._ 1973 in @Harper2007 and our study is most probably due to differences in forest definition and mapping methods between the 1953 aerial-photography derived map and the 1973 Landsat image derived map. Our study extends the forest-cover monitoring on a six decades period (from 1953 to 2014) while harmonizing the data from the previous studies. Our method provides cloud-free forest-cover estimates from 1990 to 2014 and corrections for the forest-cover in 1973.

```{r defor_comp, results="asis"}
tab.3 <- defor.comp
tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
kable(tab.3,col.names=col.names,align=c("l","l",rep("r",6)))
```

Table 3: **Comparison of Madagascar annual deforestation rates with previous studies on the period 1953-2014**. Annual deforestation rates are providing in ha/yr and in %/yr (second number in parenthesis). For deforestation rate in %/yr, exact same numbers as in scientific articles and reports from previous studies have been reported. The way annual deforestation rates in %/yr have been computed in these previous studies can slightly differ from one study to another but estimates always correct for the potential presences of clouds on forest maps. Deforestation rates in ha/yr have been recomputed from forest-cover estimates in Tab. 2 (except for @Harper2007 for the periods 1973-1990 and 1990-2000 for which annual deforestation rates in ha/yr were derived from numbers reported in the original publication, see methods) and do not correct for the potential presence of clouds.

```{r frag, results="asis"}
frag.df <- read.table(file="outputs/frag.txt", sep="\t", header=TRUE)
tab.4 <- frag.df
p.1 <- round(frag.df$interior[frag.df$Year==1973])
p.2 <- round(frag.df$interior[frag.df$Year==2014])
p.3 <- round(frag.df$patch[frag.df$Year==2014]+frag.df$transitional[frag.df$Year==2014])
names(tab.4) <- c("Year","Forest (ha)", "patch (%)", "transitional (%)", "edge (%)", "perforated (%)", "interior (%)", "NA (%)")
kable(tab.4)
```

Table 4: **Evolution of the forest fragmentation from 1950s to 2014 in Madagascar**. Forest fragmentation has progressively increased since 1953. The percentage of forest belonging to the "interior"" category (the most intact forest) has fallen from `r p.1`% in 1973 to `r p.2`% in 2014. More than `r p.3`% of the forest belong to the "patch" and "transitional" categories (isolated small forest patches) in 2014.

## 6. Figures

![Figure 1: **Ecoregions and forest types in Madagascar.** Madagascar can be divided into four climatic ecoregions with four forest types: the moist forest in the East (green), the dry forest in the West (orange), the spiny forest in the South (red), and the mangroves on the West coast (blue). Ecoregions were defined following climatic (see @Cornet1974) and vegetation (see @IEFN1996) criteria. The dark grey areas represent the remaining natural forest cover for the year 2014.](outputs/ecoregion.png)

![Figure 2: **Forest-cover change on six decades from _c._ 1953 to 2014 in Madagascar.** Forest cover changes from _c._ 1973 to 2014 are shown in the main figure, and forest cover in _c._ 1953 is shown in the bottom-right inset. Two zooms in the western dry (left part) and eastern moist (right part) ecoregions present more detailed views of (from top to bottom): forest-cover in 1950s, forest-cover change from _c._ 1973 to 2014, forest fragmentation in 2014 and distance to forest edge in 2014. Data on water bodies (blue) and water seasonality (light blue for seasonal water to dark blue for permanent water) has been extracted from @Pekel2016.](outputs/fig_fcc_highres.png)

![Figure 3: **Evolution of the mean distance to forest edge from 1953 to 2014 in Madagascar.** Black dots represent the mean distance to forest edge for each year. Vertical dashed segments represent the 95% quantiles of the distance to forest edge. We observed a continuous decrease of the mean distance to forest edge from 1953 to 2014. The mean distance was of ~1km in 1973 while it was of ~300m in 2014. Moreover, many areas were located far from forest edge (>2km and up to 4km) in 1973, while the majority of the forest was at a distance <1.2 km of the forest edge in 2014.](outputs/dist.png)

## Acknowledgements

This study is part of the Cirad's BioSceneMada project (https://bioscenemada.cirad.fr) and the Joint Research Center's ReCaREDD project (http://forobs.jrc.ec.europa.eu/recaredd). The BioSceneMada project is funded by FRB (Fondation pour la Recherche sur la Biodiversité) and the FFEM (Fond Français pour l'Environnement Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD project is funded by the European Commission.

## Data availability statement

All the data and codes used for this study are made publicly available in the `deforestmap` GitHub repository (https://github.com/ghislainv/deforestmap.git). The results and the manuscript are fully reproducible running the R script `deforestmap.R` located inside the `deforestmap` repository.

## References