### 2017-06-07 latex/deforestmap.tex
Latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Preambule %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper, 12pt, leqno]{article} %leqno: numéro d'équation à gauche
\pagenumbering{arabic} % choose how to number the pages
\usepackage{a4wide}
\usepackage[utf8]{inputenc} % accents interprétés
\usepackage{graphicx}
%\usepackage{subfig}
%\usepackage[hmargin=2cm, vmargin = 2cm, noheadfoot]{geometry} %% Pour gérer le format des pages
%\usepackage{layout} %% Pour avoir la longueur des marges
\usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
%\usepackage{multibib}
%\newcites{secnm}{Bibliographie} 
%\usepackage{verbatim} % for multiline comments
\usepackage{amssymb} %symbole de maths
\usepackage{amsmath} %idem
%\usepackage{stmaryrd} %% Symbole flèche à l'envers
%\usepackage{amsfonts}
\usepackage[english]{babel} %% Les titres en anglais
\usepackage{array} %% Pour centrer verticalement le contenu d'un tableau, entre autres...
\setcounter{secnumdepth}{4} %% Profondeur des sections, subsections
\usepackage{setspace} %% Gère l'interligne: singlespacing, doublespacing
\usepackage{booktabs}
%\singlespacing
\usepackage{longtable}
\setlength{\LTleft}{-5cm plus 1 fill}
\setlength{\LTright}{-5cm plus 1 fill}
\usepackage[colorlinks=true,citecolor=blue]{hyperref} %% Gère les hyperliens
\usepackage{lineno} %% Numérotation des lignes
\usepackage{caption2}
%\linenumbers
\newcommand{\logit}{\text{logit}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\p}{\text{p}}
\newcommand{\R}{\textnormal{\sffamily\bfseries R}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Combining global tree cover loss data with historical national
  forest-cover maps to look at six decades of deforestation and forest
  fragmentation in Madagascar}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\vspace{-1cm}
\begin{center}
%%{\large
  Ghislain Vieilledent$^{1,2,\star}$\hspace{1cm}
  Clovis Grinand$^{3}$ \\
  \vspace{0.25cm}
  Fety A. Rakotomalala$^{3}$ \hspace{1cm}
  Rija Ranaivosoa$^{4}$ \hspace{1cm}
  Jean-Roger Rakotoarijaona$^{4}$ \\
  \vspace{0.25cm}
  Thomas F. Allnutt$^{5,6}$\hspace{1cm}and\hspace{1cm}
  Frédéric Achard$^{1}$
%%}
\end{center}

\vspace{0.3cm}

{\small
  \begin{flushleft}  
    $[1]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
    $[2]$ \textbf{Cirad} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
    $[3]$ \textbf{ETC Terra}, F-75020 Paris, FRANCE\\
    $[4]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
    $[5]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
    $[6]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
    ~\\
    $[\star]$ \textbf{Corresponding author:}
    \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
    \textbackslash{Phone}:~+39.033.278.3516\\
  \end{flushleft}}

\vspace{0.3cm}

\begin{center}
  \textbf{Running headline:} \emph{Six decades of deforestation in Madagascar}
\end{center}

\newpage

\linenumbers
\doublespacing

\newcommand{\keywords}[1]{\par\noindent
{\small{\em Keywords\/}: #1}}

\begin{abstract}

  1. The island of Madagascar has an unparalleled biodiversity, mainly
  located in the tropical forests of the island, which is highly
  threatened by anthropogenic deforestation. Scattered forest maps
  from past studies at national level with substantial gaps (due to
  presence of cloud cover on satellite imagery) prevent the analyzis
  of long-term deforestation trends in Madagascar.

  2. In this study, we propose a new approach combining historical
  (1953-2000) national forest-cover maps with recent (2000-2014)
  global annual tree cover loss data to look at six decades
  (1953-2014) of deforestation and forest fragmentation in
  Madagascar. We produced new forest-cover maps at 30m-resolution over
  the full territory of Madagascar for the year 1990, and annually
  from 2000 to 2014.

  3. We estimated that Madagascar has lost 44\% of its natural forest
  cover over the period 1953-2014 (including 37\% over the period
  1973-2014).  Natural forests cover 8.9 Mha in 2014 (15\% of the
  national territory) which are divided into 4.4 Mha (50\%) of moist
  forests, 2.6 Mha (29\%) of dry forests, 1.7 Mha of spiny forests
  (19\%) and 177,000 ha (2\%) of mangroves. Since 2005, the annual
  deforestation rate has progressively increased in Madagascar to
  reach 99,000 ha/yr during 2010-2014 (corresponding to a rate of
  1.08\%/yr). Around half of the forest (46\%) is now located at less
  than 100m from the forest edge.

  4. \emph{Policy implications}: Increase in the deforestation rate
  since 2005 can be related to rapid population growth (close to
  3\%/yr) and to poor law enforcement due to political instability in
  the country.  More effort, including conservation and development
  programs, is needed in Madagascar to protect the remaining natural
  tropical forests, both to enhance local people's livelihoods and to
  support global environmental efforts.

\vspace{0.5cm}

\keywords{biodiversity, climate-change, deforestation, Landsat,
  Madagascar, tropical forest}

\end{abstract}

\newpage

\section{Introduction}
\label{introduction}

Separated from the African continent and the Indian plate about 165
and 88 million years ago respectively \citep{Ali2008}, the flora and
fauna of Madagascar followed its own evolutionary path. Isolation
combined with a high number of micro-habitats \citep{Pearson2009} has
led to Madagascar's exceptional biodiversity both in term of number of
species and endemism in many taxonomic groups \citep{Crottini2012,
  Goodman2005}.  Most of the biodiversity in Madagascar is
concentrated in the tropical forests of the island which can be
divided into three types: the moist forest in the East, the dry forest
in the West and the spiny forest in the South
\citep{Vieilledent2016}. This unparalleled biodiversity is severely
threatened by anthropogenic deforestation \citep{Harper2007,
  Vieilledent2013} associated with activities such as slash-and-burn
agriculture and pasture \citep{Scales2011}. Tropical forests in
Madagascar also store a large amount of carbon \citep{Vieilledent2016}
and high rates of deforestation in Madagascar are responsible for
large CO2 emissions in the atmosphere
\citep{Achard2014}. Deforestation threatens species survival by
directly reducing their available habitat \citep{Brooks2002,
  Tidd2001}. Forest fragmentation can also lead to species extinction
by isolating populations from each other and creating forest patches
too small to maintain viable populations
\citep{Saunders1991}. Fragmentation also increases forest edge where
ecological conditions (such as air temperature, light intensity and
air moisture) can be dramatically modified, with consequences on the
abundance and distribution of species \citep{Murcia1995}. Forest
fragmentation can also have substantial effects on forest carbon
storage capacity, as carbon stocks are much lower at the forest edge
than under a closed canopy \citep{Brinck2017}.  Moreover, forest
carbon stocks vary spatially due to climate or soil factors
\citep{Saatchi2011, Vieilledent2016}. As a consequence, accurate and
spatially explicit maps of forest-cover and forest-cover change are
necessary to monitor biodiversity loss and carbon emissions from
deforestation and forest fragmentation, assess the efficiency of
present conservation strategies \citep{Eklund2016}, and implement new
strategies for the future \citep{Vieilledent2013,
  Vieilledent2016}. Simple time-series of forest-cover estimates, such
as those provided by the FAO Forest Resource Assessment report
\citep{Keenan2015} are not sufficient.

Unfortunately, accurate and exhaustive forest-cover maps are not
available for Madagascar for the last fifteen years (2000-2015).
\citet{Harper2007} produced maps of forest cover and forest cover
changes over Madagascar for the years c. 1953, c. 1973, 1990 and 2000.
The c. 1953 forest map was derived from the visual interpretation of
aerial photography at coarse scale (1/1,000,000). Forest maps for the
years c. 1973, 1990 and 2000, were obtained from supervised
classification of Landsat satellite images at 60m resolution (for the
year 1973) or 30m resolution (for years 1990 and 2000) and can be used
to derive accurate estimates (89.5\% accuracy reported for the forest
/ non-forest map of year 2000). Maps provided by \citet{Harper2007}
were not exhaustive (due to the presence of clouds in the satellite
imagery), e.g.~11 244 km2 are mapped as unknown cover type for the
year 2000.  Using a similar supervised classification approach as in
\citet{Harper2007}, more recent maps have been produced for the
periods 2000-2005-2010 by national institutions, with the technical
support of international environmental NGOs \citep{MEFT2009,
  ONE2013}. Another set of recent forest-cover maps using an advanced
statistical tool for classification, the Random Forest classifier
\citep{Grinand2013, Rakotomalala2015}, was produced for the periods
2005-2010-2013 \citep{ONE2015}. However, these maps are either too old
to give recent estimates of deforestation \citep{MEFT2009, ONE2013},
include large areas of missing information (due to images with more
than 20\% of clouds) for the 2010 map provided by \citep{ONE2013}, or
show large mis-classification in specific areas, especially in the dry
and spiny forest domain for which the spectral answer has a strong
seasonal behavior due to the deciduousness of such forests (overall
accuracy is lower than 0.8 for the dry and spiny forests for the maps
produced by \citet{ONE2015}). Moreover, the production of such forest
maps from a supervised classification approach requires significant
resources, especially regarding the image selection step (required to
minimize cloud cover) and the training step (visual interpretation of
a large number of polygons needed to train the classification
algorithm) \citep{Rakotomalala2015}. Most of this work of image
selection and visual interpretation would need to be repeated to
produce new forest maps in the future using a similar approach.

Global forest or tree cover products have also been published recently
and can be tested at the national scale for
Madagascar. \citep{Kim2014} produced a global forest-cover change map
from 1990 to 2000 (derived from Landsat imagery). This product was
updated to cover the period 1975-2005
(\url{http://glcf.umd.edu/data/landsatFCC/}) but forest-cover maps
after 2005 were not produced. Moreover, the approach used in
\citet{Kim2014} did not accurately map the forests in the dry and
spiny ecosystems of Madagascar (see Fig. 8 in \citet{Kim2014}).
\citet{Hansen2013} mapped tree cover percentage, annual forest loss
and forest gain from 2000 to 2012 at global scale at 30 m
resolution. This product has since been updated and is now available
up to the year 2014 \citep{Hansen2013}. To map forest cover from the
\citet{Hansen2013} product, a tree cover threshold must be selected
(that defines forest cover). Selecting such a threshold is not
straightforward as the accuracy of the global tree cover map strongly
varies between forest types, and is substantially lower for dry
forests than for moist forests \citep{Bastin2017}. Moreover, the
\citet{Hansen2013} product does not provide information on
land-use. In particular the global tree cover map does not separate
tree plantations such as oil palm or eucalyptus plantations from
natural forests \citep{Tropek2014}. Thus, the global tree cover map
from \citet{Hansen2013} cannot be used alone to produce a map of
forest cover \citep{Tyukavina2017}. In complement to the tree cover
percentage provided in \citet{Hansen2013}, a layer of annual tree
cover loss is also provided (i.e.~complete loss of tree cover from a
value higher than 10\% to zero) for the period 2001-2014.

In this study, we present a simple approach which combines the maps
from \citet{Harper2007} and products from \citet{Hansen2013} to derive
annual wall-to-wall forest-cover change maps over the period 2000-2014
for Madagascar. We use the forest-cover map provided by
\citet{Harper2007} for the year 2000 (defining the land-use) with the
tree cover loss product provided by \citet{Hansen2013} that we apply
only inside forest areas identified by \citet{Harper2007}. Similar to
the approach of \citet{Harper2007}, we also assess trends in
deforestation rates and forest fragmentation from c. 1953 to 2014. The
approach described in this study can help assess the effectiveness of
current conservation strategies, and assist the implementation of
future strategies. Our approach could be easily extended to other
tropical countries that have at least one forest-cover map between
2000 and 2014. This approach can easily be repeated in the future when
the \citet{Hansen2013} products are updated.

\section{Materials and Methods}
\label{materials-and-methods}

\subsection{Creation of new forest-cover maps of Madagascar from
1953 to 2014}

We produced annual forest/non-forest maps at 30m resolution for the
full territory of Madagascar for the period 2000-2014 by combining the
forest map of year 2000 from \citet{Harper2007}, and the tree cover
percentage and annual forest cover loss maps over the period 2000-2014
from \citet{Hansen2013}. The 2000 Harper's forest map includes 208,000
ha of unclassified areas due to the presence of clouds in input
satellite imagery, mostly (88\%) within the moist forest domain which
covers 4.17 million ha in total. To provide a label (forest or
non-forest) to these unclassified pixels, we use the 2000 tree cover
percentage map of \citet{Hansen2013} by selecting a threshold of 75\%
tree cover to define forest cover as recommended by other studies for
the moist domain \citep{Achard2014}. We thus obtained a forest-cover
map for the year 2000 covering the full territory of Madagascar. We
then combined this forest-cover map of the year 2000 with the annual
forest cover loss maps from 2001 to 2014 provided by
\citet{Hansen2013} to create annual forest-cover maps from 2001 to
2014 at 30m resolution. We also completed the Harper's forest map of
year 1990 by filling unclassified areas (due to the presence of clouds
on satellite images) using our forest-cover map of year 2000. To do
so, we assumed that if forest was present in 2000, the pixel was also
forested in 1990. The remaining unclassified pixels were limited to a
relatively small total area of c. 8,000 ha. We labeled these residual
pixels as non-forest as for the year 2000.  Similarly we completed the
Harper forest map of year 1973 by filling unclassified areas using our
forest-cover map of the year 1990 assuming that if forest was present
in 1990, it was also present in 1973.  Contrary to the year 1990, the
remaining unclassified pixels for year 1973 corresponded to a
significant total area of 3.32 million ha. We also reprojected the
forest-cover map of year 1953 to a common projection in order to
compare to the other eras. This map was produced by scanning a paper
map derived from aerial photos, and thus could not be perfectly
aligned with the other maps produced through digital processing of
satellite imagery \citep{Harper2007}. Finally for all forest-cover
maps from 1973, the isolated single non-forest pixels (i.e.~fully
surrounded by forest pixels) were removed, assuming that single
non-forest pixels inside a forest patch were not corresponding to
deforestation (they might correspond to selective logging activities).
This allowed us to avoid counting very small scale events
(\textless{}0.1 ha such as selective logging) as forest fragmentation.
All the resulting maps are freely available at
\url{https://bioscenemada.cirad.fr/forestmaps}.

\subsection{Computing forest-cover areas and deforestation
rates}

From these new forest-cover maps, we calculated the total forest-cover
area for seven available years (1953-1973-1990-2000-2010-2005-2014),
and the annual deforested area and annual deforestation rate for the
corresponding six time periods between 1953 and 2014. The annual
deforestation rates were calculated as follows \citep{Puyravaud2003,
  Vieilledent2013}:

\begin{equation}
  \theta = 100 \times [1-(1-(F_{t_2}-F_{t_1})/F_{t_1})^{(1/(t_2-t_1))}
\end{equation}

where $\theta$ is the annual deforestation rate (in \%/yr),
$F_{t_2}$ and $F_{t_1}$ are the forest cover free of clouds at both
dates $t_2$ and $t_1$, and $t_2-t_1$ is the time-interval (in
years) between the two dates.

Because of the large unclassified area (3.32 million ha) in 1973, the
annual deforestation areas and rates for the two periods 1953-1973 and
1973-1990 are only indicative estimates. For these two periods the
annual deforestation rates are computed as the ratio
$(F_{t_2}-F_{t_1})/F_{t_1}$ considering only the mapped forest pixels.
Area and rate estimates are produced at the national scale and for the
four forest ecosystems present in Madagascar: moist forest in the
East, dry forest in the West, spiny forest in the South, and mangroves
on the Western coast (Fig.~\ref{fig:ecoregion}). To define the forest
domains, we used a map from the MEFT (\emph{``Ministère de
  l'Environnement et des Forêts à Madagascar''}) with the boundaries
of the four ecoregions in Madagascar.  Ecoregions were defined on the
basis of climatic and vegetation criteria using the climate
classification by \citet{Cornet1974} and the vegetation classification
from the 1996 IEFN national forest inventory \citep{IEFN1996}. Because
mangrove forests are highly dynamic ecosystems that can expand or
contract on decadal scales depending on changes in environmental
factors \citep{Armitage2015}, a fixed delimitation of the mangrove
ecoregion on six decades might not be fully appropriate. As a
consequence, our estimates of the forest-cover and deforestation rates
for mangroves in Madagascar must be considered with this limitation.

\subsection{Comparing our forest-cover and deforestation rate
estimates with previous studies}

We compared our estimates of forest-cover and deforestation rates with
estimates from the three existing studies at the national scale for
Madagascar: (i) \citep{Harper2007}, (ii) \citep{MEFT2009} and (iii)
\citep{ONE2015}. \citet{Harper2007} provides forest-cover and
deforestation estimates for the periods
c. 1953-c. 1973-1990-2000. MEFT, USAID, and CI (2009) provides
estimates for the periods 1990-2000-2005 and ONE, DGF, MNP, WCS, and
Etc Terra (2015) provides estimates for the periods 2005-2010-2013. To
compare our forest-cover and deforestation estimates over the same
time periods, we consider an additional time-period in our study
(2010-2013) by creating an extra forest-cover map for the year
2013. We computed the Pearson's correlation coefficient and the root
mean square error (RMSE) between our forest-cover estimates and
forest-cover estimates from previous studies for all the dates and
forest types (including also the total forest cover estimates). For
previous studies, the computation of annual deforestation rates (in
\%/yr) is not always detailed and might slightly differ from one study
to another \citep[see][]{Puyravaud2003}. \citet{Harper2007} also
provide total deforested areas for the two periods 1973-1990 and
1990-2000. We converted these values into annual deforested area
estimates. When annual deforested areas were not reported (for
1953-1973 in \citet{Harper2007} and in \citet{MEFT2009} and
\citet{ONE2015}), we computed them from the forest-cover estimates in
each study. These estimates cannot be corrected from the potential
bias due to the presence of residual clouds. Forest-cover and
deforestation rates were then compared between all studies for the
whole of Madagascar and the four ecoregions. The same ecoregion
boundaries as in our study were used in \citet{ONE2015} but this was
not the case for \citet{Harper2007} and \citet{MEFT2009}, which can
explain part of the differences between the estimates.

\subsection{Fragmentation}

We also conducted an analysis of changes in forest fragmentation for
the years 1953, 1973, 1990, 2000, 2010 and 2014. We applied the method
developed by \citet{Riitters2000} which uses a moving window to
characterize the fragmentation around each forested
pixel. Computations were done using the function \texttt{r.forestfrag}
of the GRASS GIS software \citep{Neteler2008}. Six categories of
fragmentation were identified from the amount of forest and its
occurrence as adjacent forest pixels: ``interior'', ``perforated'',
``edge'', ``transitional'', ``patch'', and ``undetermined''. We used a
moving window of 7x7 pixels (4.4 ha). Using this window size, forest
edge had a width of about 90m \citep{Riitters2000}. The ``interior''
category can be interpreted as the most intact forest
\citep{Potapov2017}. The ``patch'' and ``transitional'' categories
correspond to isolated small forest patches.  We reported the area of
forest in each fragmentation category for the six years and analyzed
the dynamics of fragmentation over the six decades. We also computed
the distance to forest edge for all forest pixels for the years 1953,
1973, 1990, 2000, 2005, 2010 and 2014. For that, we used the function
\texttt{gdal\_proximity.py} of the GDAL software
(\url{http://www.gdal.org/}). We computed the mean and 95\% quantiles
of the distance to forest edge and looked at the evolution of these
values with time.

\section{Results}
\label{results}

\subsection{Dynamics of forest cover and deforestation
intensity}

Natural forests in Madagascar covered 16.0 Mha in 1953, about 27\% of
the national territory of 587,041 km2. In 2014, the forest cover
dropped to 8.9 Mha, corresponding to about 15\% of the national
territory (Fig.~\ref{fig:fcc} and
Tab.~\ref{tab:forest_cover}). Madagascar has lost 44\% and 37\% of its
natural forests between 1953 and 2013, and between c. 1973 and 2014
respectively (Fig.~\ref{fig:fcc} and Tab.~\ref{tab:forest_cover}). In
2014 the remaining 8.9 Mha of natural forest were distributed as: 4.4
Mha of moist forest (50\% of total forest cover), 2.6 Mha of dry
forest (29\%), 1.7 Mha of spiny forest (19\%) and 0.18 Mha (2\%) of
mangrove forest (Fig.~\ref{fig:ecoregion} and
Tab.~\ref{tab:comp_forest}). Regarding the deforestation trend, we
observed a progressive decrease of the deforestation rate after1990
from 205,000 ha/yr (1.63\%/yr) over the period 1973-1990 to 44,300
ha/yr (0.43\%/yr) over the period 2000-2005
(Tab.~\ref{tab:forest_cover}). Then from 2005, the deforestation rate
has progressively increased and has more than doubled over the period
2010-2014 (98,700 ha/yr, 1.08\%/yr) compared to 2000-2005
(Tab.~\ref{tab:forest_cover}). The deforestation trend characterized
by a progressive decrease of the deforestation rate over the period
1990-2005 and a progressive increase of the deforestation after 2005
is valid for all four ecoregions (Tab.~\ref{tab:comp_defor}), with the
exception of the spiny forest domain for which the deforestation rate
during the period 2010-2013 was lower than during 2005-2010
(Tab.~\ref{tab:comp_defor}).

\subsection{Comparison with previous forest-cover change studies
in Madagascar}

Forest-cover maps provided by previous studies over Madagascar were
not exhaustive (unclassified areas) due to the presence of clouds in
satellite imagery used to produce such maps. In \citet{Harper2007},
the maps of years 1990 and 2000 include 0.5 and 1.12 Mha of unknown
cover type respectively. Proportions of unclassified areas are not
reported in the two other existing studies by \citet{MEFT2009} and
\citet{ONE2015}.  With our approach, we produced wall to wall
forest-cover change maps from 1990 to 2014 for the full territory of
Madagascar (Tab.~\ref{tab:forest_cover}). This allowed us to produce
more robust estimates of forest-cover and deforestation rates over
this period. Our forest-cover estimates over the period 1953-2013
(considering forest cover estimates at national level and by
ecoregions for the 6 available dates) were well correlated (0.99) to
estimates from the three previous studies (Tab.~\ref{tab:comp_forest})
with a RMSE of 300,000 ha (6\% of the mean forest cover of 4.8 Mha
when considering all dates and forest types together). These small
differences can be partly attributed to differences in ecoregion
boundaries. Despite significant differences in deforestation estimates
(Tab.~\ref{tab:comp_defor}), a similar deforestation trend was
observed across studies with a decrease of deforestation rates over
the period 1990-2005, followed by a progressive increase of the
deforestation after 2005.

\subsection{Evolution of forest fragmentation with time}

In parallel to the dynamics of deforestation, forest fragmentation has
progressively increased since 1953 in Madagascar. We observed a
continuous decrease of the mean distance to forest edge from 1953 to
2014 in Madagascar. The mean distance to forest edge has decreased to
c.  300 m in 2014 while it was previously c. 1.5 km in 1973
(Fig.~\ref{fig:dist_edge}).  Moreover, a large proportion (73\%) of
the forest was located at a distance greater than 100 m in 1973, while
almost half of the forest (46\%) was at a distance lower than 100 m
from forest edge in 2014 (Fig.  3). The percentage of forest that can
be considered intact in Madagascar has continuously decreased since
1953. The percentage of forest belonging to the ``interior'' category
(most intact forests) has fallen from 68\% in 1973 to 50\% in 2014. In
2014, more than 16\% of the forest belonged to the ``patch'' and
``transitional'' categories (isolated small forest patches) compared
to 9.5\% in 1973 (Tab.~\ref{tab:frag}).

\section{Discussion}
\label{discussion}

\subsection{Benefits of the combined use of recent global annual
  tree cover loss data with historical national forest-cover maps}

In this study, we combined recent (2001-2014) global annual tree cover
loss data \citep{Hansen2013} with historical (1953-2000) national
forest-cover maps \citep{Harper2007} to look at six decades (1953-2014)
of deforestation and forest fragmentation in Madagascar. We produced
annual forest-cover maps at 30m-resolution covering Madagascar for the
period 2000 to 2014. Our study extends the forest-cover monitoring on a
six decades period (from 1953 to 2014) while harmonizing the data from
the previous studies. We propose a generic approach to solve the problem
of forest definition which is needed to transform the 2000 global tree
cover dataset from Hansen et al. (2013) into a forest/non-forest map
\citep{Tropek2014}. We propose to use a historical national forest-cover
map, based on a national forest definition, as a forest cover mask. This
approach could be easily extended to other regions or countries for
which an accurate forest-cover map is available at any date within the
period 2000-2014, but preferably at the beginning of the period to
profit from the full record and derive long-term estimates of
deforestation. Moreover, this approach can be repeated in the future if
and when the global tree cover product is updated. We have made the
R/GRASS code used for this study freely available in a GitHub repository
(see Data availability statement) to facilitate application to other
study areas.

The accuracy of the derived forest-cover change maps depends directly on
the accuracies of the historical forest-cover maps and the tree cover
loss dataset. The reported global accuracy of the tree cover loss
dataset is 99.6\% (see Tab. S5 in \citet{Hansen2013}).
\citet{Verhegghen2016} have compared deforestation estimates derived
from the global tree cover loss dataset \citep{Hansen2013} with results
derived from semi-automated supervised classification of Landsat
satellite images \citep{Achard2014} for six countries in Central Africa
and they found a good agreement between these two sets of estimates.
Consistent with \citet{Harper2007}, we did not consider potential forest
regrowth in Madagascar (although \citet{Hansen2013} provided a tree
cover gains layer for the period 2001-2013) for the following reasons:
(i) the tree gain layer of \citet{Hansen2013} includes and catches more
easily tree plantations than natural forest regrowth \citep{Tropek2014};
(ii) there is little evidence of natural forest regeneration in
Madagascar \citep[\citet{Grouzis2001}]{Harper2007} - this can be
explained by several ecological processes following burning practice
such as soil erosion \citep{Grinand2017} and reduced seed bank due to
fire and soil loss \citep{Grouzis2001}, (iii) in areas where forest
regeneration is ecologically possible, young forest regrowth are more
easily re-burnt for agriculture and pasture; and (iv) young secondary
forests provide more limited ecosystem services compared to old-growth
natural forests in terms of biodiversity and carbon storage.

\subsection{Dynamics of forest-cover in Madagascar from 1953 to 2014}

We estimated that natural forests in Madagascar cover 8.9 Mha in 2014
(corresponding to 15\% of the country) and that Madagascar has lost 44\%
of its natural forest since 1953 (37\% since 1973). There is ongoing
scientific debate about the extent of the ``original'' forest cover in
Madagascar, and the extent to which humans have altered the natural
forest landscapes since their large-scale settlement around 800 CE
\citep{Burns2016, Cox2012}. Early French naturalists stated that the
full island was originally covered by forest
\citep{Humbert1927, Perrier1921}, leading to the common statement that
90\% of the natural forests have disappeared since the arrival of humans
on the island \citep{Kull2000}. More recent studies counter-balanced
that point of view saying that extensive areas of grassland existed in
Madagascar long before human arrival and were determined by climate,
natural grazing and other natural factors
\citep[\citet{Vorontsova2017}]{Virah-Sawmy2009}. Other authors have
questioned the entire narrative of extensive alteration of the landscape
by early human activity which, through legislation, has severe
consequences on local people \citep{Klein2002, Kull2000}. Whatever the
original proportion of natural forests and grasslands in Madagascar, our
results demonstrate that human activities since the 1950s have
profoundly impacted the natural tropical forests and that recent
conservation and development programs in Madagascar have failed to stop
deforestation in the recent years. Deforestation has strong consequences
on biodiversity and carbon emissions in Madagascar. Around 90\% of
Madagascar's species are forest dependent
\citep{Allnutt2008, Goodman2005} and \citet{Allnutt2008} estimated that
deforestation between 1953 and 2000 led to an extinction of 9\% of the
species. The additional deforestation we observed over the period
2000-2014 (around 1Mha of natural forest) worsen this result. Regarding
carbon emissions, using the 2010 aboveground forest carbon map by
\citet{Vieilledent2016}, we estimated that deforestation on the period
2010-2014 has led to 40.2 Mt C of carbon emissions in the atmosphere (10
Mt C /yr) and that the remaining aboveground forest carbon stock in 2014
is 832.8 Mt C. Associated to deforestation, we showed that the remaining
forests of Madagascar are highly fragmented with 46\% of the forest
being at less than 100m of the forest edge. Small forest fragments do
not allow to maintain viable populations and ``edge effects'' at
forest/non-forest interfaces have impacts on both carbon emissions
\citep{Brinck2017} and biodiversity loss \citep{Gibson2013, Murcia1995}.

\subsection{Deforestation trend and causes of the acceleration of the
  deforestation since 2005}

In our study, we have shown that the progressive decrease of the
deforestation rate on the period 1990-2005 was followed by a continuous
increase in the deforestation rate on the period 2005-2014. In
particular, we showed that deforestation rate has more than doubled on
the period 2010-2014 compared to 2000-2005. Our results are confirmed by
previous studies \citep{Harper2007, MEFT2009, ONE2015} despite
differences in the methodologies regarding (i) forest definition
(associated to independent visual interpretations of observation
polygons to train the classifier), (ii) classification algorithms, (iii)
deforestation rate computation method, and (iv) correction for the
presence of clouds. Our deforestation rate estimates from 1990 to 2014
have been computed from wall to wall maps at 30 m resolution and can be
considered more accurate in comparison with estimates from these
previous studies. Our forest-cover and deforestation rate estimates can
be used as source of information for the next FAO Forest Resources
Assessment project \citep{Keenan2015}. Current rates of deforestation
can also be used to build reference scenarios for deforestation in
Madagascar and contribute to the implementation of deforestation
mitigation activities in the framework of REDD+ \citep{Olander2008}.

The increase of deforestation rates after 2005 can be explained by
population growth and political instability in the country. Nearly 90\%
of Madagascar's population relies on biomass for their daily energy
needs \citep{Minten2013} and the link between population size and
deforestation has previously been demonstrated in Madagascar
\citep{Vieilledent2013, Gorenflo2011}. With a mean demographic growth
rate of about 2.8\%/yr and a population which has increased from 16 to
24 million people on the period 2000-2015 \citep{UN2015}, the increasing
demand in wood-fuel and space for agriculture is likely to explain the
increase in deforestation rates. The political crisis of 2009
\citep{Ploch2012}, followed by several years of political instability
and weak governance could also explain the increase in the deforestation
rate observed on the period 2005-2014 (Smith et al., 2003). These
results show that despite the conservation policy in Madagascar
\citep{Freudenberger2010}, deforestation has dramatically increased at
the national level since 2005. More efforts should be done to identify
the real causes and mechanisms explaining this rampant deforestation in
order to implement new conservation strategies that will be effective at
conserving the few remaining natural tropical forests in Madagascar.

\newpage

\section{Author's contribution}
\label{authors-contribution}

All authors conceived the ideas and designed methodology; GV analysed
the data and wrote the {\R}/GRASS script; GV drafted the manuscript. All
authors contributed critically to the drafts and gave final approval for
publication.

\section{Acknowledgements}
\label{acknowledgements}

This study is part of the Cirad's BioSceneMada project
(\url{https://bioscenemada.cirad.fr}) and the Joint Research Center's
ReCaREDD project (\url{http://forobs.jrc.ec.europa.eu/recaredd}). The
BioSceneMada project is funded by FRB (Fondation pour la Recherche sur
la Biodiversité) and the FFEM (Fond Français pour l'Environnement
Mondial) under the project agreement AAP-SCEN-2013 I. The ReCaREDD
project is funded by the European Commission.

\section{Data accessibility}
\label{data-accessibility}

All the data and codes used for this study are made publicly available
in the \texttt{deforestmap} GitHub repository
(\url{https://github.com/ghislainv/deforestmap.git}). The results and
the manuscript are fully reproducible running the {\R} script
\texttt{deforestmap.R} located inside the \texttt{deforestmap}
repository.

\newpage
\singlespacing

\bibliographystyle{gcb}
\bibliography{bib/biblio}

\newpage

\section{Tables}
\label{tables}

\nopagebreak

\vfill
\begin{table}[!h]
  \begin{longtable}[]{@{}rrrrr@{}}
    \toprule
    Year & Forest (ha) & Unmap (ha) & Annual defor. (ha/yr) & Rate
    (\%/yr)\tabularnewline
    \midrule
    \endhead
    1953 & 15968176 & 0 & - & -\tabularnewline
    1973 & 14242592 & 3316531 & 86279 & 0.57\tabularnewline
    1990 & 10762442 & 0 & 204715 & 1.63\tabularnewline
    2000 & 9879031 & 0 & 88341 & 0.85\tabularnewline
    2005 & 9667553 & 0 & 42296 & 0.43\tabularnewline
    2010 & 9319851 & 0 & 69540 & 0.73\tabularnewline
    2014 & 8925246 & 0 & 98651 & 1.08\tabularnewline
    \bottomrule
  \end{longtable}
  \addtocounter{table}{-1}

  \caption{\textbf{Evolution of the forest cover and deforestation rates
      from 1953 to 2014 in Madagascar}. Forest map for the year 1973 has 3.3
    Mha of unclassified areas due to the presence of clouds on satellite
    images. As a consequence, deforestation rates for the periods 1953-1973
    and 1973-1990 are indicative. The two last columns indicate the
    deforestation rates on the previous time-period (e.g.~1953-1973 for year
    1973, 1973-1990 for year 1990, etc.).}

  \label{tab:forest_cover}
\end{table}
\vfill

\newpage

\vspace*{\stretch{1}}
\begin{table}[!h]
  {\footnotesize
  \begin{longtable}[]{@{}llrrrrrrrr@{}}
    \toprule
    Forest type & Source & 1953 & 1973 & 1990 & 2000 & 2005 & 2010 & 2013 &
    2014\tabularnewline
    \midrule
    \endhead
    Total & Harper2007 & 15995900 & 14173100 & 10605700 & 8982100 & - & - &
    - & -\tabularnewline
    & MEFT2009 & - & - & 10650142 & 9678402 & 9413218 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 9451350 & 8977337 & 8485509 &
    -\tabularnewline
    & this study & 15968176 & 14242592 & 10762494 & 9879031 & 9667553 &
    9319851 & 9051029 & 8925246\tabularnewline
    Moist & Harper2007 & 8765600 & 6876000 & 5234300 & 4166800 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 5270599 & 4787771 & 4700430 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 4555788 & 4457184 & 4345093 &
    -\tabularnewline
    & this study & 8578299 & 6989942 & 5270169 & 4872016 & 4767876 & 4633104
    & 4470194 & 4409842\tabularnewline
    Dry & Harper2007 & 4252100 & 4027700 & 2711800 & 2457000 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 3320582 & 3084976 & 3027505 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 3223028 & 2970192 & 2678640 &
    -\tabularnewline
    & this study & 4761551 & 4434871 & 3224917 & 2940970 & 2880819 & 2734639
    & 2642253 & 2595621\tabularnewline
    Spiny & Harper2007 & 2978200 & 3029800 & 2420000 & 2132200 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 2123630 & 1871735 & 1756884 & - & - &
    -\tabularnewline
    & ONE2015 & - & - & - & - & 1681527 & 1558533 & 1466765 &
    -\tabularnewline
    & this study & 2462830 & 2582880 & 2054724 & 1857628 & 1810704 & 1744427
    & 1731308 & 1712731\tabularnewline
    Mangroves & Harper2007 & - & - & 239600 & 226100 & - & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & - & - & - & - & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 173564 & 171220 & 169877 & -\tabularnewline
    & this study & 143412 & 199853 & 181226 & 177708 & 177492 & 177149 &
    176890 & 176718\tabularnewline
    \bottomrule
  \end{longtable}}
  \addtocounter{table}{-1}

  \caption{\textbf{Comparing our estimates of forest-cover (in ha) for
      Madagascar with previous studies on the period 1953-2014}. We compared
    our forest-cover with the estimates from three previous studies
    \citep{Harper2007, MEFT2009, ONE2015}. We computed a The increase in
    mangrove and spiny forest covers from \emph{c.} 1953 to \emph{c.} 1973
    in \citet{Harper2007} and our study is most probably due to differences
    in forest definition and mapping methods between the 1953
    aerial-photography derived map and the 1973 Landsat image derived map.}

  \label{tab:comp_forest}
\end{table}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{table}[!h]
  {\footnotesize
  \begin{longtable}[]{@{}llrrrrrr@{}}
    \toprule
    Forest type & Source & 1953-1973 & 1973-1990 & 1990-2000 & 2000-2005 &
    2005-2010 & 2010-2013\tabularnewline
    \midrule
    \endhead
    Total & Harper2007 & 91140 (0.30) & 200206 (1.70) & 80740 (0.90) & - & -
    & -\tabularnewline
    & MEFT2009 & - & - & 97174 (0.83) & 53037 (0.53) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 94803 (1.18) & 163943 (1.50)\tabularnewline
    & this study & 86279 (0.57) & 204712 (1.63) & 88346 (0.85) & 42296
    (0.43) & 69540 (0.73) & 89607 (0.97)\tabularnewline
    Moist & Harper2007 & 94480 (0.60) & 87188 (1.70) & 32200 (0.80) & - & -
    & -\tabularnewline
    & MEFT2009 & - & - & 48283 (0.79) & 17468 (0.35) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 19721 (0.50) & 37364 (0.94)\tabularnewline
    & this study & 79418 (1.02) & 101163 (1.65) & 39815 (0.78) & 20828
    (0.43) & 26954 (0.57) & 54303 (1.19)\tabularnewline
    Dry & Harper2007 & 11220 (0.20) & 77153 (1.90) & 19820 (0.70) & - & - &
    -\tabularnewline
    & MEFT2009 & - & - & 23561 (0.67) & 11494 (0.40) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 50567 (1.80) & 97184 (2.29)\tabularnewline
    & this study & 16334 (0.35) & 71174 (1.86) & 28395 (0.92) & 12030 (0.41)
    & 29236 (1.04) & 30795 (1.14)\tabularnewline
    Spiny & Harper2007 & -2580 (-0.10) & 35865 (1.20) & 28170 (1.20) & - & -
    & -\tabularnewline
    & MEFT2009 & - & - & 25190 (1.19) & 22970 (1.23) & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 24599 (1.69) & 30589 (1.66)\tabularnewline
    & this study & -6002 (-0.24) & 31068 (1.34) & 19710 (1.00) & 9385 (0.51)
    & 13255 (0.74) & 4373 (0.25)\tabularnewline
    Mangroves & Harper2007 & - & - & 550 (0.20) & - & - & -\tabularnewline
    & MEFT2009 & - & - & - & - & - & -\tabularnewline
    & ONE2015 & - & - & - & - & 469 (0.32) & 448 (0.20)\tabularnewline
    & this study & -2822 (-1.67) & 1096 (0.57) & 352 (0.20) & 43 (0.02) & 69
    (0.04) & 86 (0.05)\tabularnewline
    \bottomrule
  \end{longtable}}
  \addtocounter{table}{-1}

  \caption{\textbf{Comparing our estimates of annual deforestation
      rates for Madagascar with previous studies on the period
      1953-2014}. Annual deforestation areas (in ha/yr) and annual
    deforestation rates (second number in parenthesis, in \%/yr) are
    provided. For deforestation rates in \%/yr, exact same numbers as
    in scientific articles and reports from previous studies have been
    reported. The way annual deforestation rates in \%/yr have been
    computed in these previous studies can slightly differ from one
    study to another but estimates always correct for the potential
    presences of clouds on forest maps. Annual deforested areas in
    ha/yr have been recomputed from forest-cover estimates in
    Tab.~\ref{tab:comp_forest} (except for \citet{Harper2007} for the
    periods 1973-1990 and 1990-2000 for which annual deforested ares
    in ha/yr were derived from numbers reported in the original
    publication, see methods) and do not correct for the potential
    presence of clouds.}

  \label{tab:comp_defor}
\end{table}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{table}[!h]

  {\small
  \begin{longtable}[]{@{}rrrrrrrr@{}}
    \toprule
    Year & Forest (ha) & patch (\%) & transitional (\%) & edge (\%) &
    perforated (\%) & interior (\%) & NA (\%)\tabularnewline
    \midrule
    \endhead
    1953 & 15962870 & 0.01 & 1.12 & 4.46 & 0.58 & 93.83 &
    0.00\tabularnewline
    1973 & 14228217 & 2.21 & 7.25 & 19.81 & 2.86 & 67.87 &
    0.01\tabularnewline
    1990 & 10749572 & 3.00 & 8.17 & 21.28 & 3.81 & 63.73 &
    0.01\tabularnewline
    2000 & 9866145 & 3.09 & 8.37 & 22.13 & 3.92 & 62.49 &
    0.01\tabularnewline
    2005 & 9659861 & 3.51 & 8.88 & 22.56 & 6.44 & 58.59 &
    0.02\tabularnewline
    2010 & 9306528 & 4.28 & 9.72 & 22.94 & 8.52 & 54.52 &
    0.02\tabularnewline
    2014 & 8911481 & 5.18 & 10.72 & 23.25 & 10.58 & 50.24 &
    0.03\tabularnewline
    \bottomrule
  \end{longtable}}
  \addtocounter{table}{-1}

  \caption{\textbf{Evolution of the forest fragmentation from 1953 to 2014
      in Madagascar}. Six categories of fragmentation were identified from the
    amount of forest and its occurrence as adjacent forest pixels:
    ``interior'', ``perforated'', ``edge'', ``transitional'', ``patch'', and
    ``undetermined'' \citep{Riitters2000}. We used a moving window of 7x7
    pixels (4.4 ha). Using this window size, forest edge had a width of
    about 90m. The ``interior'' category can be interpreted as the most
    intact forest. The ``patch'' and ``transitional'' categories correspond
    to isolated small forest patches.}

  \label{tab:frag}
\end{table}
\vspace*{\stretch{1}}

\newpage

\section{Figures}
\label{figures}

\nopagebreak

\vfill
\begin{figure}[h!]
  \centering

    \includegraphics[width=11cm]{outputs/ecoregion.png}

    \caption{\textbf{Ecoregions and forest types in Madagascar.}
      Madagascar can be divided into four climatic ecoregions with four forest
      types: the moist forest in the East (green), the dry forest in the West
      (orange), the spiny forest in the South (red), and the mangroves on the
      West coast (blue). Ecoregions were defined following climatic
      \citep{Cornet1974} and vegetation \citep{IEFN1996} criteria. The dark
      grey areas represent the remaining natural forest cover for the year
      2014.}

    \label{fig:ecoregion}

\end{figure}
\vfill

\newpage

\vspace*{\stretch{1}}
\begin{figure}[h!]
  \centering

  \includegraphics[width=\textwidth]{outputs/fig_fcc_highres.png}

  \caption{\textbf{Forest-cover change on six decades from 1953
      to 2014 in Madagascar.} Forest cover changes from \emph{c.} 1973 to 2014
    are shown in the main figure, and forest cover in \emph{c.} 1953 is
    shown in the bottom-right inset. Two zooms in the western dry (left
    part) and eastern moist (right part) ecoregions present more detailed
    views of (from top to bottom): forest-cover in 1950s, forest-cover
    change from \emph{c.} 1973 to 2014, forest fragmentation in 2014 and
    distance to forest edge in 2014. Data on water bodies (blue) and water
    seasonality (light blue for seasonal water to dark blue for permanent
    water) has been extracted from \citet{Pekel2016}.}

  \label{fig:fcc}

\end{figure}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{figure}[h!]
  \centering

  \includegraphics[width=13cm]{outputs/dist.png}

  \caption{\textbf{Evolution of the distance to forest edge from
      1953 to 2014 in Madagascar.} Black dots represent the mean distance to
    forest edge for each year. Vertical dashed segments represent the 90\%
    quantiles (5\% and 95\%) of the distance to forest edge. Horizontal
    dashed grey line indicates a distance to forest edge of 100~m.
    Percentages indicate the percentage of forest at a distance to forest
    edge lower than 100~m for each year.}

  \label{fig:dist_edge}

\end{figure}
\vspace*{\stretch{1}}

\end{document}