### 2018-10-11 update_2017/update_2017.Rnw
Update 2017

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Preambule %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper, 12pt, leqno, dvipsnames]{article} %leqno: numéro d'équation à gauche
\pagenumbering{arabic} % choose how to number the pages
\usepackage{a4wide}
\usepackage[utf8]{inputenc} % accents interprétés
\usepackage{graphicx}
%\usepackage{subfig}
%\usepackage[hmargin=2cm, vmargin = 2cm, noheadfoot]{geometry} %% Pour gérer le format des pages
%\usepackage{layout} %% Pour avoir la longueur des marges
\usepackage[round,sort]{natbib} %% Natbib is a popular style for formatting references.
%\usepackage{multibib}
%\newcites{secnm}{Bibliographie} 
%\usepackage{verbatim} % for multiline comments
\usepackage{amssymb} %symbole de maths
\usepackage{amsmath} %idem
%\usepackage{stmaryrd} %% Symbole flèche à l'envers
%\usepackage{amsfonts}
\usepackage[english]{babel} %% Les titres en anglais
\usepackage{array} %% Pour centrer verticalement le contenu d'un tableau, entre autres...
\setcounter{secnumdepth}{4} %% Profondeur des sections, subsections
\usepackage{setspace} %% Gère l'interligne: singlespacing, doublespacing
\usepackage{booktabs}
%\singlespacing
\usepackage{longtable}
\usepackage[table]{xcolor}
\setlength{\LTleft}{-5cm plus 1 fill}
\setlength{\LTright}{-5cm plus 1 fill}
\usepackage[colorlinks=true,citecolor=Blue,urlcolor=Maroon]{hyperref} %% Gère les hyperliens
%\usepackage{lineno} %% Numérotation des lignes
%\linenumbers
\newcommand{\logit}{\text{logit}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\p}{\text{p}}
\newcommand{\R}{\textnormal{\sffamily\bfseries R}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

<<options, echo=FALSE, results="hide">>=
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, cache=FALSE,
               #results="hide", 
               warning=FALSE,
               message=FALSE, highlight=TRUE,
               fig.show="hide", size="small",
               fig.align="center",
               tidy=FALSE)
options(knitr.kable.NA="-")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Update to: Combining global tree cover loss data with historical national
  forest cover maps to look at six decades of deforestation and forest
  fragmentation in Madagascar$^{\S}$}
\author{}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\vspace{-1cm}
\begin{center}
  Ghislain Vieilledent$^{1,\star}$, Clovis Grinand$^{2}$,
  Fety A. Rakotomalala$^{2}$,\\
  Rija Ranaivosoa$^{3}$, Jean-Roger Rakotoarijaona$^{3}$,\\
  Thomas F. Allnutt$^{4,5}$, and Frédéric Achard$^{6}$\\
\end{center}

\vspace{0.3cm}

{\small
\noindent$^{\S}$ This article was initially published as:

\noindent\textbf{Vieilledent G., C. Grinand, F. A. Rakotomalala, R. Ranaivosoa, J.-R. Rakotoarijaona, T. F. Allnutt, and F. Achard.} 2018. Combining global tree cover loss data with historical national forest-cover maps to look at six decades of deforestation and forest fragmentation in Madagascar. \emph{Biological Conservation.} \textbf{222}: 189-197. [doi: \href{https://doi.org/10.1016/j.biocon.2018.04.008}{10.1016/j.biocon.2018.04.008}].}

\vspace{0.3cm}

{\small
  \begin{flushleft}  
    $[1]$ \textbf{CIRAD} -- UPR Forêts et Sociétés, F-34398 Montpellier, FRANCE\\
    $[2]$ \textbf{Nitid\ae}, F-75020 Paris, FRANCE\\
    $[3]$ \textbf{Office National pour l'Environnement}, 101 Antananarivo, MADAGASCAR\\
    $[4]$ \textbf{Wildlife Conservation Society}, 101 Antananarivo, MADAGASCAR\\
    $[5]$ \textbf{GreenInfo Network}, Oakland, California, USA\\
    $[6]$ \textbf{Joint Research Center of the European Commission} -- Bio-economy Unit, I-21027 Ispra (VA), ITALY\\
    ~\\
    $[\star]$ \textbf{Corresponding author:}
    \textbackslash{E-mail}:~ghislain.vieilledent@cirad.fr
  \end{flushleft}}

% \vspace{0.3cm}

% {\small
%   \begin{flushleft}
%     \textbf{Highlights:}\\
%     - We produced new 30 m resolution forest cover maps for Madagascar for the period 2000-2014.\\
%     - Madagascar has lost 44\% of its natural forest cover over the period 1953-2014.\\
%     - Annual deforestation rate has increased in Madagascar since 2005 to reach 1.1\%/yr in 2014.\\
%     - Half of the tropical forest in Madagascar is now located at less than 100~m from forest edge.\\
%     - Conservation efforts must be intensified to save Madagascar unique forests and biodiversity.\\
%   \end{flushleft}
% }

\newpage

\doublespacing
%\linenumbers

\newcommand{\keywords}[1]{\par\noindent
{\small{\em Keywords\/}: #1}}

\begin{abstract}

Blabla

\vspace{0.5cm}

\keywords{biodiversity, climate-change, deforestation, forest-fragmentation, Madagascar, tropical forest}

\end{abstract}

\newpage

\section{Introduction}
\label{introduction}

Blabla

\section{Data accessibility}
\label{data-accessibility}

Data and code used for this study have been made permanently and publicly available
on the CIRAD Dataverse repository so that the results are entirely reproducible:
\begin{itemize}
\item Input data: \url{http://dx.doi.org/10.18167/DVN1/2FP7LR}
\item Code: \url{http://dx.doi.org/10.18167/DVN1/275TDF}
\item Output data: \url{http://dx.doi.org/10.18167/DVN1/AUBRRC}
\end{itemize}

\newpage
\singlespacing

\renewcommand{\bibname}{References}
\renewcommand{\bibsection}{\section{\bibname}}
\bibliographystyle{bib/besjournals}
\bibliography{bib/biblio}

\newpage

\section{Tables}
\label{tables}

\nopagebreak

\vfill
<<forest_cover, results="asis">>=
# Import and format
defor.df <- read.table(file="tabs/defor.txt", sep="\t", header=TRUE)
perc.1 <- round(100*(defor.df$area[defor.df$Year==1973]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1973])
perc.2 <- round(100*(defor.df$area[defor.df$Year==1953]-defor.df$area[defor.df$Year==2014])/defor.df$area[defor.df$Year==1953])
d.1 <- defor.df$theta[defor.df$Year==1990]
d.2 <- defor.df$theta[defor.df$Year==2010]
tab.1 <- defor.df
tab.1[,c(2:4)] <- round(tab.1[,c(2:4)]/1000)
tab.1$Year <- as.character(tab.1$Year)
names(tab.1) <- c("Year", "Forest (Kha)", "Unmap (Kha)", "Annual defor. (Kha/yr)", "Rate (%/yr)")
# Caption
tabcap1 <- "\\textbf{Change in natural forest cover and deforestation rates from 1953 to 2017 in Madagascar.} Areas are provided in thousands of hectares (Kha). Forest map for the year 1973 has 3.3 Mha of unclassified areas due to the presence of clouds on satellite images. As a consequence, partial deforestation rates for the periods 1953-1973 and 1973-1990 are computed based on the available forest extent. The last two columns indicate the annual deforested areas and annual deforestation rates on the previous time-period (e.g.~1953-1973 for year 1973, 1973-1990 for year 1990, etc.)."
# Kable
knitr::kable(tab.1, digits=1, format.args=list(big.mark=","), booktabs=TRUE, format="latex", 
             caption=tabcap1, align=c("l","r","r","r","r"), linesep="") %>% 
  kable_styling(latex_options=c("striped","hold_position"))
@
\vfill

\newpage

\vspace*{\stretch{1}}
<<comp_forest, results="asis">>=
# Import and format
fcc.comp <- read.table(file="tabs/fcc_comp.txt", sep=" ", header=TRUE)
tab.2 <- fcc.comp
tab.2 <- tab.2[,c(2,1,3:10)]
tab.2[,c(3:10)] <- round(tab.2[,c(3:10)]/1000)
tab.2$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
col.names <- c("Forest type","Source","1953","1973","1990","2000","2005","2010","2013","2017")
# Caption
tabcap2 <- "\\textbf{Comparing Madagascar forest cover estimates with
      previous studies on the period 1953-2017}. We compared our
    estimates of forest cover with the estimates from three previous
    studies \\citep{Harper2007, MEFT2009, ONE2015}. Areas are provided
    in thousands of hectares (Kha). We obtained a Pearson's
    correlation coefficient of 0.99 between our forest cover estimates
    and forest cover estimates from previous studies. The increase in
    mangrove and spiny forest covers from 1953 to 1973 in \\citet{Harper2007} and our study is most probably due to
    differences in forest definition and mapping methods between the
    1953 aerial-photography derived map and the 1973 Landsat image
    derived map."
# Kable
knitr::kable(tab.2, col.names=col.names, align=c("l","l",rep("r",8)), format.args=list(big.mark=","), 
             booktabs=TRUE, format="latex", caption=tabcap2, linesep=c("","","","\\addlinespace")) %>% 
  kable_styling(latex_options=c("striped","hold_position","scale_down"))
@
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
<<comp_defor, results="asis">>=
# Import and format
tab.3 <- read.table(file="tabs/defor_comp.txt", sep=" ", header=TRUE)
tab.3$ForestType <- c("Total",rep("",3),"Moist",rep(" ",3),"Dry",rep(" ",3),"Spiny",rep(" ",3),"Mangroves",rep(" ",3))
col.names <- c("Forest type","Source","1953-1973","1973-1990","1990-2000","2000-2005","2005-2010","2010-2013")
# Caption
tabcap3 <- "\\textbf{Comparing Madagascar annual deforestation rates
      with previous studies on the period 1953-2013}. Annual
    deforested areas (in thousands of hectares per year, Kha/yr) and
    annual deforestation rates (second number in parenthesis, in
    \\%/yr) are provided. For deforestation rates in \\%/yr, exact same
    numbers as in scientific articles and reports from previous
    studies \\citep{Harper2007, MEFT2009, ONE2015} have been
    reported. The way annual deforestation rates in \\%/yr have been
    computed in these previous studies can slightly differ from one
    study to another, but estimates always correct for the potential
    presences of clouds on satellite images and unclassified areas on
    forest maps. Annual deforested areas in Kha/yr have been
    recomputed from forest cover estimates in
    Tab.~\\ref{tab:comp_forest} (except for \\citet{Harper2007} for the
    periods 1973-1990 and 1990-2000 for which annual deforested areas
    in Kha/yr were derived from numbers reported in the original
    publication, see methods) and do not correct for the potential
    presence of clouds."
# Kable
knitr::kable(tab.3, col.names=col.names, align=c("l","l",rep("r",6)), booktabs=TRUE, format="latex",
             caption=tabcap3, linesep=c("","","","\\addlinespace")) %>% 
  kable_styling(latex_options=c("striped","hold_position","scale_down"))
@
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
<<frag, results="asis">>=
# Import and format
frag.df <- read.table(file="tabs/fordens.txt", sep="\t", header=TRUE)
tab.4 <- frag.df
tab.4$Year <- as.character(tab.4$Year)
tab.4$forest <- round(tab.4$forest/1000)
names(tab.4) <- c("Year","Forest (Kha)", "0-20", "21-40", "41-60", "61-80", "81-100")
# Caption
tabcap4 <- "\\textbf{Change in forest fragmentation from 1953 to
      2017 in Madagascar}. Five forest fragmentation classes, based on
    the percentage of forest in the neighborhood, are defined: 0-20\\%
    (highly fragmented), 21-40\\%, 41-60\\%, 61-80\\% and 81-100\\% (lowly
    fragmented). The percentage of forest falling in each forest
    fragmentation class is reported for each year. Forest areas are
    provided in thousands of hectares (Kha)."
# Kable
knitr::kable(tab.4, digits=0, format.args=list(big.mark=","), booktabs=TRUE, format="latex",
             caption=tabcap4, linesep="") %>%
  kable_styling(latex_options=c("striped","hold_position"))
@
\vspace*{\stretch{1}}

\newpage

\section{Figures}
\label{figures}

\nopagebreak

\vfill
\begin{figure}[h!]
  \centering

    \includegraphics[width=10cm]{figs/ecoregion.png}

    \caption{\textbf{Ecoregions and forest types in Madagascar.}
      Madagascar can be divided into four climatic ecoregions with
      four forest types: the moist forest in the East (green), the dry
      forest in the West (orange), the spiny forest in the South
      (red), and the mangroves on the West coast (blue). Ecoregions
      were defined following climatic \citep{Cornet1974} and
      vegetation \citep{IEFN1996} criteria. The dark grey areas
      represent the remaining natural forest cover for the year 2017.
      Forest types are defined on the basis of their belonging to
      one of the four ecoregions.}

    \label{fig:ecoregion}

\end{figure}
\vfill

\newpage

\vspace*{\stretch{1}}
\begin{figure}[h!]
  \centering

  \includegraphics[width=16cm]{figs/fig_fcc.png}

  \caption{\textbf{Forest cover change on six decades from 1953 to
      2017 in Madagascar.} forest cover changes from 1973 to
    2017 are shown in the main figure, and forest cover in 1953 is shown 
    in the bottom-right inset. Two zooms in the western
    dry (left part) and eastern moist (right part) ecoregions present
    more detailed views of (from top to bottom): forest cover in
    1953, forest cover change from 1973 to 2017, forest
    fragmentation in 2017 and distance to forest edge in 2017. Data on
    water bodies (blue) and water seasonality (light blue for seasonal
    water to dark blue for permanent water) has been extracted from
    \citet{Pekel2016}.}

  \label{fig:fcc}

\end{figure}
\vspace*{\stretch{1}}

\newpage

\vspace*{\stretch{1}}
\begin{figure}[h!]
  \centering

  \includegraphics[width=12cm]{figs/dist.png}

  \caption{\textbf{Change in distance to forest edge from 1953
      to 2017 in Madagascar.} Black dots represent the mean distance
    to forest edge for each year. Vertical dashed segments represent
    the 90\% quantiles (5\% and 95\%) of the distance to forest
    edge. Horizontal dashed grey line indicates a distance to forest
    edge of 100~m. Numbers at the bottom of each vertical segments are
    the percentage of forest at a distance to forest edge lower than
    100~m for each year.}

  \label{fig:dist_edge}

\end{figure}
\vspace*{\stretch{1}}

\end{document}